---
title: "Sugarpint Incubation 12/01/2020"
author: "Nick Framsted"
date: "12/8/2020"
output: pdf_document
---
# Summary:
This script contains analysis of data generated from a preliminary incubation of rocks collected from sugarpoint state park on 11/29/20 as well as a comparison of PAR data from the field on 07/20/20 and PAR data collected from this lab incubation.

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# importing and cleaning up data
```{r}
dat <- read_csv("data/201201_sugarpt_lab_inc_R.csv", skip = 1)
str(dat)
#View(dat)

# formatting date times column to posixct
dat$date_time_PST <- dmy_hms(dat$date, tz = "America/Los_Angeles")
head(dat)

# filtering out rows with error codes, and adding in chamber numbers based on sensor column
dat_2 <- dat %>%
  filter(is.na(error)) %>%
  mutate(chamber = ifelse(sensor == "Batch 3 (PSt3)", "Chamber 1", ifelse(sensor == "Batch 3_1 (PSt3)", "Chamber 2", ifelse(sensor == "Batch 3_2 (PSt3)", "Chamber 3", "Chamber 4"))))
head(dat_2)

# selecting only columns I need
dat_subset <- dat_2 %>%
  select(date, delta_T_min, oxygen, temp, chamber, date_time_PST)

```

# plotting O2 signals
```{r}
# plotting only data points with no error code
dat_subset %>%
  ggplot(aes(x = delta_T_min, y = oxygen)) +
  geom_point() +
  facet_wrap(~chamber)

# plot of all data points (including rows with errors)
#first need to add chamber column to dat dataframe
dat_1 <- dat %>%
  mutate(chamber = ifelse(sensor == "Batch 3 (PSt3)", "Chamber 1", ifelse(sensor == "Batch 3_1 (PSt3)", "Chamber 2", ifelse(sensor == "Batch 3_2 (PSt3)", "Chamber 3", "Chamber 4")))) %>%
  filter(oxygen < 10.5 & oxygen > 0) #filtering out outlier datapoints

o2_plot <- dat_1 %>%
  ggplot(aes(x = delta_T_min, y = oxygen, color = error)) +
  geom_point() +
  scale_color_viridis_d(na.value = "grey") +
  facet_wrap(~chamber) +
  theme_bw(base_size = 20) +
  theme(legend.position = "right", legend.title = element_text(size = 12), legend.text = element_text(size = 10))
o2_plot

ggsave("plots/201201_sugarpt_inc/error_plot.png", plot = o2_plot, width = 10, height = 7)
# lots of crazy data points in the data with errors included
```

# calculating metabolic rate 
```{r}
# adding regression lines
o2_plot +
  geom_smooth(method = "lm") +
  stat_regline_equation()

# adding column in data based on light and dark periods to use for grouping of data points
head(dat_1)
dat_1 <- dat_1 %>%
  mutate(photoperiod = ifelse(date_time_PST >= "2020-12-01 17:08:00", "dark", "light"))

metab_plot <- dat_1 %>%
  ggplot(aes(x = delta_T_min, y = oxygen, group = photoperiod)) +
  geom_point(aes(color = photoperiod)) +
  geom_smooth(method = "lm") +
  facet_wrap(~chamber) +
  scale_color_viridis_d()
metab_plot

# still some spurious values dark photoperiod first started
# filtering data to remove some points in that transition from light to dark
dat_1_clean <- dat_1 %>%
  filter(date_time_PST < "2020-12-01 17:05:00" | date_time_PST > "2020-12-01 17:15:00")

# replotting using cleaned data
plot_metab_clean <- dat_1_clean %>%
   ggplot(aes(x = delta_T_min, y = oxygen, group = photoperiod)) +
  geom_point(aes(color = photoperiod)) +
  geom_smooth(method = "lm") +
  facet_wrap(~chamber) +
  scale_color_viridis_d(name = "Photoperiod", labels = c("Dark", "Light")) +
  labs(y = expression(Oxygen~Concentration~(mg~L^{-1})), x = "Elapsed Time (min)") +
  theme_bw(base_size = 15)
plot_metab_clean

ggsave("plots/201201_sugarpt_inc/metab_clean.png", plot = plot_metab_clean, width = 10, height = 7)

# adding equations to plots
plot_metab_trendline <- plot_metab_clean +
  stat_regline_equation(label.x.npc = c(0.5, 0), label.y.npc = c(0.65, 1))

ggsave("plots/201201_sugarpt_inc/metab_trendline.png", plot = plot_metab_trendline, width = 10, height = 7)

```

# plotting field PAR data
```{r}
# reading in PAR data from tahoe in-situ incubations
# field light level data
PAR_field <- read_csv("data/R_concat_2020_07_20_180000Z.txt", skip = 2)
head(PAR_field)

# converting time into posixct
PAR_field$date_time_PST <- as_datetime(PAR_field$Time_sec, tz = "America/Los_Angeles")
head(PAR_field)

# comparing light levels in the field to replicated levels in the lab to verify if they are simiar

PAR_field %>%
  ggplot(aes(x = date_time_PST, y = PAR_umol_s_m_2)) +
  geom_line()

# viewing only pineland data
PAR_field %>%
  filter(date_time_PST < "2020-07-20 11:03:00") %>%
  ggplot(aes(x = date_time_PST, y = PAR_umol_s_m_2)) +
  geom_line()

# viewing only sugarpine st park data
PAR_field %>%
  filter(date_time_PST > "2020-07-20 16:00:00") %>%
  ggplot(aes(x = date_time_PST, y = PAR_umol_s_m_2)) +
  geom_line()

```

According to the field PAR data, light levels at pineland ~11AM were ~1800umol m^-2 s^-1, and at sugarpine state park around ~4:20PM they ranged from 600-800. 

Average light levels in Stan Loeb's thesis on periphyton range 1457 umol m^-2 s^-1, this was a rough back-of-the-envelope conversion from W m^-2 to umol m^-2 s^-1.

# plotting lab PAR data
```{r}
# reading in lab PAR data
PAR_lab <- read_csv("data/R_2020-12-02_PARdat.txt", skip = 2)

# converting date times to posixct
PAR_lab$date_time_PST <- as_datetime(PAR_lab$Time_sec, tz = "America/Los_Angeles")
head(PAR_lab)

# plotting lab data
PAR_lab %>%
  ggplot(aes(x = date_time_PST, y = PAR_umol_s_m_2)) +
  geom_line()

# filtering out unnecessary data
PAR_lab %>%
  filter(date_time_PST < "2020-12-01 17:07:00") %>%
  ggplot(aes(x = date_time_PST, y = PAR_umol_s_m_2)) +
  geom_line()
```

