---
title: "210613_pnld_A_E_analysis"
author: "Nick Framsted"
date: "7/26/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# Importing and Processing data
```{r, include = FALSE}
# importing and cleaning up metabolism data from both ambient and enriched incubations
dat_a <- read_csv("data/210613_pineland_a_e/210613_pnld_ambient_rates.csv")

# selecting only the columns that I need
dat_a <- dat_a %>%
  dplyr::select(chamber, total_DW_g, total_AFDW_g, rock_SA_m2, NEP_mg_d:nutrient_trt)

dat_e <- read_csv("data/210613_pineland_a_e/210613_pnld_enriched_rates.csv")

# selecting only the columns that I need
dat_e <- dat_e %>%
  dplyr::select(chamber, total_DW_g, total_AFDW_g, rock_SA_m2, NEP_mg_d:nutrient_trt)


# binding the rows of the two dataframes together
dat <- bind_rows(dat_a, dat_e)
# adding in nutrient enriched dummy variable column (0 = ambient, 1 = enriched nutrients)
dat <- dat %>%
  mutate(nutrients = ifelse(nutrient_trt == "ambient", 0, 1))

# making ER rates positive so that we can log-transform them in subsequent models
dat <- dat %>%
  mutate(ER_mg_d = -1 * ER_mg_d, ER_mgO2_d_gAFDW = -1 * ER_mgO2_d_gAFDW, ER_mgO2_d_m2 = -1 * ER_mgO2_d_m2)

write_csv(dat, "data/main_analysis/210613_pnld_A_E_rates.csv")

#View(dat)

```

# AFDW-normalized Metabolic rates
### GPP vs temperature and nutrients
```{r}
# linear model of GPP and temp
GPP_AFDW_mod <- lm(log(GPP_mgO2_d_gAFDW) ~ temp_C*nutrients, dat)
summary(GPP_AFDW_mod)

# checking that model assumptions are met
plot(GPP_AFDW_mod)


dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

# simplified plot
GPP_AFDW_model_plot <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

GPP_AFDW_model_plot

ggsave("plots/210613_pineland_ambient_enriched/210613_GPP_AFDW_model_plot.png", plot = GPP_AFDW_model_plot, width = 7, height = 3.5)

# dataset with single outlier (chamber 1) removed
dat_clean <- dat %>%
  filter(!chamber == 1)

# model of ER with single outlier (chamber 1) removed
GPP_AFDW_mod2 <- lm(log(GPP_mgO2_d_gAFDW) ~ temp_C*nutrients, dat_clean)
summary(GPP_AFDW_mod2)

plot(GPP_AFDW_mod2)

dat_clean %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))
```

### ER vs temperature and nutrients
```{r}
ER_AFDW_mod <- lm(log(ER_mgO2_d_gAFDW) ~ temp_C*nutrients, dat)
summary(ER_AFDW_mod)

plot(ER_AFDW_mod)

# interpreting effect size of temperature effect
exp(coef(ER_AFDW_mod)["temp_C"])
# looks that GPP increases about 10% for every 1C temp increase

dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~gAFDW^{-1})))


# simplified plot w/o axis labels for use in plotly--This is used to identify which chambers correspond to outlier datapoints
p <- dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(title = "ER response to Temperature")

ER_plot <- plotly::ggplotly(p)
ER_plot


# model of ER with single outlier (chamber 1) removed
ER_AFDW_mod2 <- lm(log(ER_mgO2_d_gAFDW) ~ temp_C*nutrients, dat_clean)
summary(ER_AFDW_mod2)

# Plot of this model
dat_clean %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~gAFDW^{-1})))

```
Chamber 1 had unusually large ER, though it's not quite clear why. This chamber had the lowest AFDW of all chambers (not by a large margin), and an average-looking total DW. The rock was average-sized according to the water volume measurements for the chambers.

According to the first model (with chamber 1 included), temp had a significant effect on ER, with ER increasing by 11% for each 1C temperature increase.

### NEP vs. temp and nutrients
```{r}
# linear model of NEP and temp
NEP_AFDW_mod <- lm(log(NEP_mgO2_d_gAFDW) ~ temp_C*nutrients, dat)
summary(NEP_AFDW_mod)

plot(NEP_AFDW_mod)

dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, color = nutrient_trt)) +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~gAFDW)))

# simplified plot w/o axis labels for use in plotly--This is used to identify which chambers correspond to outlier datapoints
p <- dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, color = nutrient_trt)) +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(title = "NEP response to Temperature")

NEP_plot <- plotly::ggplotly(p)
NEP_plot

# dataset with single outlier (chamber 1) removed
dat_clean <- dat %>%
  filter(!chamber == 1)

# model of ER without potential outlier
NEP_AFDW_mod2 <- lm(NEP_mgO2_d_gAFDW ~ temp_C*nutrients, dat_clean)
summary(NEP_AFDW_mod2)

dat_clean %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, color = nutrient_trt)) +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~gAFDW)))
```
Looks like chamber 1 is an outlier and has unusually high NEP in the 23.5C treatment. Again, it's not clear what is the cause of this outlier besides measurement error.

# SA-normalized metabolic rates
### GPP vs temperature and nutrients model
```{r}
# linear model of GPP and temp
GPP_SA_mod <- lm(log(GPP_mgO2_d_m2) ~ temp_C*nutrients, dat)
summary(GPP_SA_mod)

plot(GPP_SA_mod)

# interpreting effect size of temperature on GPP
exp(coef(GPP_SA_mod)["temp_C"])
# looks that GPP increases about 6% for every 1C temp increase

GPP_SA_plot <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "GPP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~m^{-2})))
GPP_SA_plot

# making simplified plot w/o expressions for use in plotly function
GPP_SA <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic()
# interactive plotly plot to get info on individual data points
GPP_SA_plotly <- plotly::ggplotly(GPP_SA)
GPP_SA_plotly
```
Temp was found to have a significant effect on SA-normalized GPP, with a 1C temperature increase corresponding to a 6% increase in GPP.

### ER vs temp and nutrients
```{r}
# linear model of GPP and temp
ER_SA_mod <- lm(log(ER_mgO2_d_m2) ~ temp_C*nutrients, dat)
summary(ER_SA_mod)

plot(ER_SA_mod)

# interpreting effect size of temperature effect
exp(coef(ER_SA_mod)["temp_C"])
# looks that ER increases about 9.8% for every 1C temp increase

dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~m^{-2})))

# making simplified plot without expressions in the axis labels for plotly
ER_SA_plot <- dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic()

# ploting interactive plot in plotly to identify which datapoints correspond to which chambers
ER_SA_plotly <- plotly::ggplotly(ER_SA_plot)
ER_SA_plotly

```
Looks like temp has a significant effect on ER, with a 1C increase corresponding to a 9.8% increase in ER. Nutrients and the temp:nutrient interaction effects each were nearly significant with p-values of 0.054 and 0.094 respectively.

### NEP vs temp and nutrients
```{r}
# linear model of NEP and temp
NEP_SA_mod <- lm(log(NEP_mgO2_d_m2) ~ temp_C*nutrients, dat)
summary(NEP_SA_mod)

plot(NEP_SA_mod)

# interpreting effect size of temperature effect
exp(coef(NEP_SA_mod)["temp_C"])
# looks that ER increases about 5% for every 1C temp increase

dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~m^{-2})))

# making simplified plot without expressions in the axis labels for plotly
NEP_SA_plot <- dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_m2, color = nutrient_trt)) +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  theme_classic()

# ploting interactive plot in plotly to identify which datapoints correspond to which chambers
NEP_SA_plotly <- plotly::ggplotly(NEP_SA_plot)
NEP_SA_plotly
```
Temp had a significant effect on NEP, increasing it by 5% for every 1C temperature increase.

# Plotting additional Variables to asses outlier datapoints
```{r}
#View(dat)
# plot of total AFDW of all chambers
dat %>%
  ggplot(aes(x = chamber, y = total_AFDW_g)) +
  geom_col() +
  theme_classic()

# plot of total DW of all chambers
dat %>%
  ggplot(aes(x = chamber, y = total_DW_g)) +
  geom_col() +
  theme_classic()

# plot of rock surface area of all chambers
dat %>%
  ggplot(aes(x = chamber, y = rock_SA_m2)) +
  geom_col() +
  theme_classic()

# calculating average total AFDW for thi incubation to compare to that of other incubations
mean(dat$total_AFDW_g)


### calculating average AFDW/m2 for the ambient and enriched incubations
avg_AFDW <- dat %>%
  filter(nutrient_trt == "ambient") %>% # subsetting to only include ambient since same rocks were used for ambient and enriched and would artificially double the sample size used to calculate the sd
  summarize(avg_AFDW_g_m2 = mean(total_AFDW_g/rock_SA_m2), sd_AFDW_g_m2 = sd(total_AFDW_g/rock_SA_m2))
avg_AFDW

```
Looks like average AFDW/m2 during the april 2020 incubations was 23 +- 8 g/m2, which is nearly double that of this incubation. Thus there may be a lack of temperature and nutrient effects either due to this, or due to a lack of temperature response at this higher range of temperatures.

# Double Logarithmic graphs (as specified in section 7.3.3 of del Giorgio & Williams 2005)

## GPP
```{r}
# linear model of log-transformed GPP (biomass-normalized) and log-transformed temperature

log_GPP_AFDW_mod <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat)
summary(log_GPP_AFDW_mod)

plot(log_GPP_AFDW_mod)
###### Model interpretations ######

# no significant predictors in this model

################ Removing Outliers ###########

# dataset with single outlier (chamber 1) removed
dat_clean <- dat %>%
  filter(!chamber == 1)

# model of ER with single outlier (chamber 1) removed
log_GPP_AFDW_mod2 <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat_clean)
summary(log_GPP_AFDW_mod2)

plot(log_GPP_AFDW_mod2)

GPP_log_log <- dat_clean %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #stat_regline_equation() +
  #scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))


ggsave("plots/210613_pineland_ambient_enriched/210613_GPP_log_log_plot.png", plot = GPP_log_log, width = 4, height = 2.5)
```

## ER
```{r}
# linear model of log-transformed ER rates (biomass normalized) and log-transformed temperature
log_ER_AFDW_mod <- lm(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat)
summary(log_ER_AFDW_mod)

plot(log_ER_AFDW_mod)

###### Model interpretations ######

# interpreting effect size of temperature effect
exp(coef(log_ER_AFDW_mod)["log(temp_C)"])


ER_log_log <- dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~gAFDW^{-1})))

ggsave("plots/210613_pineland_ambient_enriched/210613_ER_log_log_plot.png", plot = ER_log_log, width = 4, height = 2.5)
```

## NEP
```{r}
# linear model of log-transformed NEP (biomass normalized) and log-transformed temp
log_NEP_AFDW_mod <- lm(log(NEP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat)
summary(log_NEP_AFDW_mod)

plot(log_NEP_AFDW_mod)

# no significant predictors in the model


NEP_log_log <- dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~gAFDW)))

ggsave("plots/210613_pineland_ambient_enriched/210613_NEP_log_log_plot.png", plot = NEP_log_log, width = 4, height = 2.5)
```
