---
title: "210808_pnld_ambient"
author: "Nick Framsted"
date: "8/17/2021"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# Importing and Processing data
```{r, include = FALSE}
dat <- read_csv("data/210808_pineland_a_e/210808_pineland_ambient_O2.csv", skip = 1)
head(dat)
#View(dat)

# removing bottom row of NAs from dataframe
dat <- dat[-44745,]
#View(dat)

# Creating tank_channel column to assign chambers to individual data points in DO data by concatenating Device Serial and Channel columns. Keeping original columns as well. 
dat <- dat %>%
 unite(tank_channel, Device_Serial, Channel, sep = "_", remove = FALSE)


#View(dat)
#head(dat)



# filtering out rows with error codes, and adding in chamber numbers based the tank and channel number cross referenced against incubation notes
## double checked for correctness on 8/17/21 NTF

dat_2 <- dat %>%
  filter(Error == 0 & !is.na(Value)) %>%
  mutate(
    chamber = case_when(
    tank_channel == "Tank 1_1"  ~ "1",
    tank_channel == "Tank 1_2"  ~ "8",
    tank_channel == "Tank 1_3" ~ "11",
    tank_channel == "Tank 1_4" ~ "12",
    tank_channel == "Tank 2_1" ~ "3",
    tank_channel == "Tank 2_2" ~ "5",
    tank_channel == "Tank 2_3" ~ "7",
    tank_channel == "Tank 2_4" ~ "14",
    tank_channel == "Tank 3_1" ~ "4",
    tank_channel == "Tank 3_2" ~ "9",
    tank_channel == "Tank 3_3" ~ "13",
    tank_channel == "Tank 3_4" ~ "15",
    tank_channel == "Tank 4_1" ~ "2",
    tank_channel == "Tank 4_2" ~ "6",
    tank_channel == "Tank 4_3" ~ "10",
    tank_channel == "Tank 4_4" ~ "16",
  )) %>%
   mutate(treatment = case_when(
     Device_Serial == "Tank 1" ~ "22.5C",
     Device_Serial == "Tank 2" ~ "19.5C",
     Device_Serial == "Tank 3" ~ "25.5C",
     Device_Serial == "Tank 4" ~ "28.5C"))
 
#View(dat_2)
head(dat_2)

# renaming oxygen data column and the delta_T columns to include units
dat_2 <- dat_2 %>%
  rename(oxygen = Value, time = Time, date = Date, temp = Temp, delta_T_min = delta_t)
head(dat_2)

### Merging DO data with chamber data ###

# selecting only columns I need, and changing chamber column from a factor to a double
dat_subset <- dat_2 %>%
  select(date, time, delta_T_min, oxygen, temp, chamber, treatment)
#View(dat_subset)
  dat_subset$chamber <- as.numeric(dat_subset$chamber)

# importing in other complementary data associated with chambers
dat_chambers <- read_csv("data/210808_pineland_a_e/210808_pineland_A_E_datasheet.csv")
#str(dat_chambers)
dat_chambers$chamber <- as.numeric(dat_chambers$chamber)
#View(dat_chambers)

# joining DO data and chamber data by "chamber" column
dat_full <- full_join(dat_subset, dat_chambers, by = "chamber")
#View(dat_full)

# converting oxygen concentration (mg/L) to oxygen mass (mg) by multipying by water volumes in chambers
dat_full <- dat_full %>%
  mutate(oxygen_mass_mg = oxygen * water_volume_L)
#View(dat_full)
```


# Plotting temp and DO in chambers
```{r echo = FALSE}
# making treatment a factor variable and re-ordering it for plots
dat_full$treatment <- factor(dat_full$treatment, levels = c("19.5C", "22.5C", "25.5C", "28.5C"))
# making chamber a factor so that it's a discrete value for coloring datapoints in plot
dat_full$chamber <- as.factor(dat_full$chamber)

# plotting Do concentrations in each chamber
DO_conc_plot <- dat_full %>%
  ggplot(aes(x = delta_T_min, y = oxygen, color = chamber)) +
  geom_line(size = 1.5) +
  scale_color_viridis_d() +
  facet_wrap(~treatment) +
  labs(y = expression(Oxygen~Concentration~(mg~L^{-1})), x = "Elapsed Time (min)", title = "Pineland Ambient Incubation DO") +
  theme_bw(base_size = 20)

DO_conc_plot

ggsave("plots/210808_pineland_ambient_enriched/210808_pineland_ambient_DO_conc_plot.png", plot = DO_conc_plot, width = 10, height = 7)

# plotting temperature for each chamber
temp_plot <- dat_full %>%
  ggplot(aes(x = delta_T_min, y = temp, color = chamber)) +
  geom_line(size = 1.5) +
  scale_color_viridis_d() +
  facet_wrap(~treatment) +
  labs(y = expression('Temperature (\u00B0C)'), x = "Elapsed Time (min)", title = "Pineland Ambient Incubation Temperature") +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")

temp_plot

ggsave("plots/210808_pineland_ambient_enriched/210808_pineland_ambient_temp_plot.png", plot = temp_plot, width = 10, height = 7)

# plotting DO mass (mg) for each chamber
DO_mass_plot <- dat_full %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg, color = chamber)) +
  geom_line(size = 1.5) +
  scale_color_viridis_d() +
  facet_wrap(~treatment) +
  labs(y = expression(Oxygen~Mass~(mg)), x = "Elapsed Time (min)", title = "Pineland Ambient Incubation DO Mass") +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")

  DO_mass_plot
  
ggsave("plots/210808_pineland_ambient_enriched/210808_pineland_ambient_DO_mass_plot.png", plot = DO_mass_plot, width = 10, height = 7)

# making interactive plot with plotly to help identity data points for rate calculations below
p<- plotly::ggplotly(DO_mass_plot)
p
# exporting interactive plot to html
htmlwidgets::saveWidget(as_widget(p), "plots/210808_pineland_ambient_enriched/210808_pineland_ambient_DO_mass_plot.html")

```


## Metabolic Rate Calculations
### Tank 2; 19.5C
#### Chamber 3
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_3 <- dat_full %>%
  filter(chamber == "3" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_3_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_3)
summary(pnld_a_3_NEP_mod)
pnld_a_3_NEP_mod$coefficients[2]
pnld_a_3_NEP_slope <- unname(pnld_a_3_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_3_NEP <- pnld_a_3_NEP_slope * 1440
pnld_a_3_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_3 <- dat_full %>%
  filter(chamber =="3" & delta_T_min >= 95)

pnld_a_3_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_3)
summary(pnld_a_3_ER_mod)
pnld_a_3_ER_mod$coefficients[2]
pnld_a_3_ER_slope <- unname(pnld_a_3_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_3_ER <- pnld_a_3_ER_slope * 1440
pnld_a_3_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_3_plot <- dat_full %>%
  filter(chamber == "3") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_3_NEP_mod$coefficients[1], slope = pnld_a_3_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_3_ER_mod$coefficients[1], slope = pnld_a_3_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_3_plot
```

#### Chamber 5
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_5 <- dat_full %>%
  filter(chamber == "5" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_5_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_5)
summary(pnld_a_5_NEP_mod)
pnld_a_5_NEP_mod$coefficients[2]
pnld_a_5_NEP_slope <- unname(pnld_a_5_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_5_NEP <- pnld_a_5_NEP_slope * 1440
pnld_a_5_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_5 <- dat_full %>%
  filter(chamber =="5" & delta_T_min >= 95)

pnld_a_5_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_5)
summary(pnld_a_5_ER_mod)
pnld_a_5_ER_mod$coefficients[2]
pnld_a_5_ER_slope <- unname(pnld_a_5_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_5_ER <- pnld_a_5_ER_slope * 1440
pnld_a_5_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_5_plot <- dat_full %>%
  filter(chamber == "5") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_5_NEP_mod$coefficients[1], slope = pnld_a_5_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_5_ER_mod$coefficients[1], slope = pnld_a_5_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_5_plot
```

#### Chamber 7
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_7 <- dat_full %>%
  filter(chamber == "7" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_7_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_7)
summary(pnld_a_7_NEP_mod)
pnld_a_7_NEP_mod$coefficients[2]
pnld_a_7_NEP_slope <- unname(pnld_a_7_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_7_NEP <- pnld_a_7_NEP_slope * 1440
pnld_a_7_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_7 <- dat_full %>%
  filter(chamber =="7" & delta_T_min >= 95)

pnld_a_7_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_7)
summary(pnld_a_7_ER_mod)
pnld_a_7_ER_mod$coefficients[2]
pnld_a_7_ER_slope <- unname(pnld_a_7_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_7_ER <- pnld_a_7_ER_slope * 1440
pnld_a_7_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_7_plot <- dat_full %>%
  filter(chamber == "7") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_7_NEP_mod$coefficients[1], slope = pnld_a_7_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_7_ER_mod$coefficients[1], slope = pnld_a_7_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_7_plot
```

#### Chamber 14
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_14 <- dat_full %>%
  filter(chamber == "14" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_14_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_14)
summary(pnld_a_14_NEP_mod)
pnld_a_14_NEP_mod$coefficients[2]
pnld_a_14_NEP_slope <- unname(pnld_a_14_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_14_NEP <- pnld_a_14_NEP_slope * 1440
pnld_a_14_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_14 <- dat_full %>%
  filter(chamber =="14" & delta_T_min >= 95)

pnld_a_14_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_14)
summary(pnld_a_14_ER_mod)
pnld_a_14_ER_mod$coefficients[2]
pnld_a_14_ER_slope <- unname(pnld_a_14_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_14_ER <- pnld_a_14_ER_slope * 1440
pnld_a_14_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_14_plot <- dat_full %>%
  filter(chamber == "14") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_14_NEP_mod$coefficients[1], slope = pnld_a_14_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_14_ER_mod$coefficients[1], slope = pnld_a_14_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_14_plot
```

### Tank 1; 22.5C
#### Chamber 1
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_1 <- dat_full %>%
  filter(chamber == "1" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_1_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_1)
summary(pnld_a_1_NEP_mod)
pnld_a_1_NEP_mod$coefficients[2]
pnld_a_1_NEP_slope <- unname(pnld_a_1_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_1_NEP <- pnld_a_1_NEP_slope * 1440
pnld_a_1_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_1 <- dat_full %>%
  filter(chamber =="1" & delta_T_min >= 95)

pnld_a_1_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_1)
summary(pnld_a_1_ER_mod)
pnld_a_1_ER_mod$coefficients[2]
pnld_a_1_ER_slope <- unname(pnld_a_1_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_1_ER <- pnld_a_1_ER_slope * 1440
pnld_a_1_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_1_plot <- dat_full %>%
  filter(chamber == "1") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_1_NEP_mod$coefficients[1], slope = pnld_a_1_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_1_ER_mod$coefficients[1], slope = pnld_a_1_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_1_plot
```

#### Chamber 8
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_8 <- dat_full %>%
  filter(chamber == "8" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_8_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_8)
summary(pnld_a_8_NEP_mod)
pnld_a_8_NEP_mod$coefficients[2]
pnld_a_8_NEP_slope <- unname(pnld_a_8_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_8_NEP <- pnld_a_8_NEP_slope * 1440
pnld_a_8_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95 and before minute 200).
pnld_a_8 <- dat_full %>%
  filter(chamber =="8" & delta_T_min >= 95 & delta_T_min <= 200)

pnld_a_8_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_8)
summary(pnld_a_8_ER_mod)
pnld_a_8_ER_mod$coefficients[2]
pnld_a_8_ER_slope <- unname(pnld_a_8_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_8_ER <- pnld_a_8_ER_slope * 1440
pnld_a_8_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_8_plot <- dat_full %>%
  filter(chamber == "8") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_8_NEP_mod$coefficients[1], slope = pnld_a_8_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_8_ER_mod$coefficients[1], slope = pnld_a_8_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_8_plot
```

#### Chamber 11
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_11 <- dat_full %>%
  filter(chamber == "11" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_11_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_11)
summary(pnld_a_11_NEP_mod)
pnld_a_11_NEP_mod$coefficients[2]
pnld_a_11_NEP_slope <- unname(pnld_a_11_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_11_NEP <- pnld_a_11_NEP_slope * 1440
pnld_a_11_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_11 <- dat_full %>%
  filter(chamber =="11" & delta_T_min >= 95)

pnld_a_11_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_11)
summary(pnld_a_11_ER_mod)
pnld_a_11_ER_mod$coefficients[2]
pnld_a_11_ER_slope <- unname(pnld_a_11_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_11_ER <- pnld_a_11_ER_slope * 1440
pnld_a_11_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_11_plot <- dat_full %>%
  filter(chamber == "11") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_11_NEP_mod$coefficients[1], slope = pnld_a_11_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_11_ER_mod$coefficients[1], slope = pnld_a_11_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_11_plot
```

#### Chamber 12
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_12 <- dat_full %>%
  filter(chamber == "12" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_12_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_12)
summary(pnld_a_12_NEP_mod)
pnld_a_12_NEP_mod$coefficients[2]
pnld_a_12_NEP_slope <- unname(pnld_a_12_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_12_NEP <- pnld_a_12_NEP_slope * 1440
pnld_a_12_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_12 <- dat_full %>%
  filter(chamber =="12" & delta_T_min >= 95)

pnld_a_12_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_12)
summary(pnld_a_12_ER_mod)
pnld_a_12_ER_mod$coefficients[2]
pnld_a_12_ER_slope <- unname(pnld_a_12_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_12_ER <- pnld_a_12_ER_slope * 1440
pnld_a_12_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_12_plot <- dat_full %>%
  filter(chamber == "12") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_12_NEP_mod$coefficients[1], slope = pnld_a_12_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_12_ER_mod$coefficients[1], slope = pnld_a_12_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_12_plot
```

### Tank 3; 25.5C
#### Chamber 4
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_4 <- dat_full %>%
  filter(chamber == "4" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_4_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_4)
summary(pnld_a_4_NEP_mod)
pnld_a_4_NEP_mod$coefficients[2]
pnld_a_4_NEP_slope <- unname(pnld_a_4_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_4_NEP <- pnld_a_4_NEP_slope * 1440
pnld_a_4_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_4 <- dat_full %>%
  filter(chamber =="4" & delta_T_min >= 95)

pnld_a_4_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_4)
summary(pnld_a_4_ER_mod)
pnld_a_4_ER_mod$coefficients[2]
pnld_a_4_ER_slope <- unname(pnld_a_4_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_4_ER <- pnld_a_4_ER_slope * 1440
pnld_a_4_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_4_plot <- dat_full %>%
  filter(chamber == "4") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_4_NEP_mod$coefficients[1], slope = pnld_a_4_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_4_ER_mod$coefficients[1], slope = pnld_a_4_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_4_plot
```

#### Chamber 9
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_9 <- dat_full %>%
  filter(chamber == "9" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_9_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_9)
summary(pnld_a_9_NEP_mod)
pnld_a_9_NEP_mod$coefficients[2]
pnld_a_9_NEP_slope <- unname(pnld_a_9_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_9_NEP <- pnld_a_9_NEP_slope * 1440
pnld_a_9_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_9 <- dat_full %>%
  filter(chamber =="9" & delta_T_min >= 95)

pnld_a_9_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_9)
summary(pnld_a_9_ER_mod)
pnld_a_9_ER_mod$coefficients[2]
pnld_a_9_ER_slope <- unname(pnld_a_9_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_9_ER <- pnld_a_9_ER_slope * 1440
pnld_a_9_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_9_plot <- dat_full %>%
  filter(chamber == "9") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_9_NEP_mod$coefficients[1], slope = pnld_a_9_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_9_ER_mod$coefficients[1], slope = pnld_a_9_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_9_plot
```

#### Chamber 13
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_13 <- dat_full %>%
  filter(chamber == "13" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_13_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_13)
summary(pnld_a_13_NEP_mod)
pnld_a_13_NEP_mod$coefficients[2]
pnld_a_13_NEP_slope <- unname(pnld_a_13_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_13_NEP <- pnld_a_13_NEP_slope * 1440
pnld_a_13_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_13 <- dat_full %>%
  filter(chamber =="13" & delta_T_min >= 95)

pnld_a_13_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_13)
summary(pnld_a_13_ER_mod)
pnld_a_13_ER_mod$coefficients[2]
pnld_a_13_ER_slope <- unname(pnld_a_13_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_13_ER <- pnld_a_13_ER_slope * 1440
pnld_a_13_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_13_plot <- dat_full %>%
  filter(chamber == "13") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_13_NEP_mod$coefficients[1], slope = pnld_a_13_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_13_ER_mod$coefficients[1], slope = pnld_a_13_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_13_plot
```

#### Chamber 15
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_15 <- dat_full %>%
  filter(chamber == "15" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_15_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_15)
summary(pnld_a_15_NEP_mod)
pnld_a_15_NEP_mod$coefficients[2]
pnld_a_15_NEP_slope <- unname(pnld_a_15_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_15_NEP <- pnld_a_15_NEP_slope * 1440
pnld_a_15_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_15 <- dat_full %>%
  filter(chamber =="15" & delta_T_min >= 95)

pnld_a_15_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_15)
summary(pnld_a_15_ER_mod)
pnld_a_15_ER_mod$coefficients[2]
pnld_a_15_ER_slope <- unname(pnld_a_15_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_15_ER <- pnld_a_15_ER_slope * 1440
pnld_a_15_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_15_plot <- dat_full %>%
  filter(chamber == "15") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_15_NEP_mod$coefficients[1], slope = pnld_a_15_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_15_ER_mod$coefficients[1], slope = pnld_a_15_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_15_plot
```

### Tank 4; 28.5C
#### Chamber 2
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_2 <- dat_full %>%
  filter(chamber == "2" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_2_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_2)
summary(pnld_a_2_NEP_mod)
pnld_a_2_NEP_mod$coefficients[2]
pnld_a_2_NEP_slope <- unname(pnld_a_2_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_2_NEP <- pnld_a_2_NEP_slope * 1440
pnld_a_2_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_2 <- dat_full %>%
  filter(chamber =="2" & delta_T_min >= 95)

pnld_a_2_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_2)
summary(pnld_a_2_ER_mod)
pnld_a_2_ER_mod$coefficients[2]
pnld_a_2_ER_slope <- unname(pnld_a_2_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_2_ER <- pnld_a_2_ER_slope * 1440
pnld_a_2_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_2_plot <- dat_full %>%
  filter(chamber == "2") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_2_NEP_mod$coefficients[1], slope = pnld_a_2_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_2_ER_mod$coefficients[1], slope = pnld_a_2_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_2_plot
```

#### Chamber 6
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_6 <- dat_full %>%
  filter(chamber == "6" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_6_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_6)
summary(pnld_a_6_NEP_mod)
pnld_a_6_NEP_mod$coefficients[2]
pnld_a_6_NEP_slope <- unname(pnld_a_6_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_6_NEP <- pnld_a_6_NEP_slope * 1440
pnld_a_6_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_6 <- dat_full %>%
  filter(chamber =="6" & delta_T_min >= 95)

pnld_a_6_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_6)
summary(pnld_a_6_ER_mod)
pnld_a_6_ER_mod$coefficients[2]
pnld_a_6_ER_slope <- unname(pnld_a_6_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_6_ER <- pnld_a_6_ER_slope * 1440
pnld_a_6_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_6_plot <- dat_full %>%
  filter(chamber == "6") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_6_NEP_mod$coefficients[1], slope = pnld_a_6_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_6_ER_mod$coefficients[1], slope = pnld_a_6_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_6_plot
```

#### Chamber 10
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_10 <- dat_full %>%
  filter(chamber == "10" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_10_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_10)
summary(pnld_a_10_NEP_mod)
pnld_a_10_NEP_mod$coefficients[2]
pnld_a_10_NEP_slope <- unname(pnld_a_10_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_10_NEP <- pnld_a_10_NEP_slope * 1440
pnld_a_10_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_10 <- dat_full %>%
  filter(chamber =="10" & delta_T_min >= 95)

pnld_a_10_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_10)
summary(pnld_a_10_ER_mod)
pnld_a_10_ER_mod$coefficients[2]
pnld_a_10_ER_slope <- unname(pnld_a_10_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_10_ER <- pnld_a_10_ER_slope * 1440
pnld_a_10_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_10_plot <- dat_full %>%
  filter(chamber == "10") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_10_NEP_mod$coefficients[1], slope = pnld_a_10_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_10_ER_mod$coefficients[1], slope = pnld_a_10_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_10_plot
```

#### Chamber 16
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 21 to minute 95 ##

pnld_a_16 <- dat_full %>%
  filter(chamber == "16" & delta_T_min >= 21 & delta_T_min <= 95)

pnld_a_16_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_16)
summary(pnld_a_16_NEP_mod)
pnld_a_16_NEP_mod$coefficients[2]
pnld_a_16_NEP_slope <- unname(pnld_a_16_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_16_NEP <- pnld_a_16_NEP_slope * 1440
pnld_a_16_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 95).
pnld_a_16 <- dat_full %>%
  filter(chamber =="16" & delta_T_min >= 95)

pnld_a_16_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_16)
summary(pnld_a_16_ER_mod)
pnld_a_16_ER_mod$coefficients[2]
pnld_a_16_ER_slope <- unname(pnld_a_16_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_16_ER <- pnld_a_16_ER_slope * 1440
pnld_a_16_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_16_plot <- dat_full %>%
  filter(chamber == "16") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_16_NEP_mod$coefficients[1], slope = pnld_a_16_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_16_ER_mod$coefficients[1], slope = pnld_a_16_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_16_plot
```

## Creating dataframe with metab rates
```{r}
# creating dataframe of metab rates to make barplot
pnld_a_rates <- data.frame(
  "NEP_mg_d" = c(pnld_a_1_NEP, pnld_a_2_NEP, pnld_a_3_NEP, pnld_a_4_NEP, pnld_a_5_NEP, pnld_a_6_NEP, pnld_a_7_NEP, pnld_a_8_NEP, pnld_a_9_NEP, pnld_a_10_NEP, pnld_a_11_NEP, pnld_a_12_NEP, pnld_a_13_NEP, pnld_a_14_NEP, pnld_a_15_NEP, pnld_a_16_NEP),
  "ER_mg_d" = c(pnld_a_1_ER, pnld_a_2_ER, pnld_a_3_ER, pnld_a_4_ER, pnld_a_5_ER, pnld_a_6_ER, pnld_a_7_ER, pnld_a_8_ER, pnld_a_9_ER, pnld_a_10_ER, pnld_a_11_ER, pnld_a_12_ER, pnld_a_13_ER, pnld_a_14_ER, pnld_a_15_ER,  pnld_a_16_ER),
  "chamber" = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16"))
head(pnld_a_rates)
#str(pnld_a_rates)

# formatting data types of columns
pnld_a_rates$NEP_mg_d <- as.numeric(pnld_a_rates$NEP_mg_d)
pnld_a_rates$ER_mg_d <- as.numeric(pnld_a_rates$ER_mg_d)

# calculating GPP from NEP and ER columns
pnld_a_rates <- pnld_a_rates %>%
  mutate(GPP_mg_d = NEP_mg_d - ER_mg_d)
head(pnld_a_rates)
```

## Joining metab rates dataframe with chamber dataframe
```{r}
# changing chamber column in datasheet to character to be compatable to joining to pnld_a_rates
dat_chambers$chamber <- as.factor(dat_chambers$chamber)

# joining chamber and metabolic rate data
dat_final <- full_join(dat_chambers, pnld_a_rates, by = "chamber")
#View(dat_final)

# adding temperature treatment (discrete factor) and temp_C (continuous numeric) columns based off of tank column already in data
dat_final <- dat_final %>%
  mutate(
    treatment = case_when(
      Tank_a == "1" ~ "22.5C",
      Tank_a == "2" ~ "19.5C",
      Tank_a == "3" ~ "25.5C",
      Tank_a == "4" ~ "28.5C"
    )) %>%
  mutate(
    temp_C = case_when(
           Tank_a == "1" ~ 22.5,
           Tank_a == "2" ~ 19.5,
           Tank_a == "3" ~ 25.5,
           Tank_a == "4" ~ 28.5
           ))


#View(dat_final)
```

## Normalizing metabolic rates
```{r}
# normalizing metabolic rates by AFDW and by rock surface area
dat_final <- dat_final %>%
  mutate(GPP_mgO2_d_gAFDW = GPP_mg_d/total_AFDW_g, NEP_mgO2_d_gAFDW = NEP_mg_d/total_AFDW_g, ER_mgO2_d_gAFDW = ER_mg_d/total_AFDW_g) %>%
  mutate(GPP_mgO2_d_m2 = GPP_mg_d/rock_SA_m2, NEP_mgO2_d_m2 = NEP_mg_d/rock_SA_m2, ER_mgO2_d_m2 = ER_mg_d/rock_SA_m2)

#View(dat_final)
```

# Exporting metabolic rate data for further analysis
```{r}
# making identifier column for joining to ambient incubation data for nutrient effect analysis
dat_final <- dat_final %>%
  mutate(nutrient_trt = "ambient")
# exporting data for final analysis
write_csv(dat_final, "data/210808_pineland_a_e/210808_pnld_ambient_rates.csv")
```