---
title: "210808_pnld_A_E_analysis"
author: "Nick Framsted"
date: "8/17/2021"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# Importing and Processing data
```{r, include = FALSE}
# importing and cleaning up metabolism data from both ambient and enriched incubations
dat_a <- read_csv("data/210808_pineland_a_e/210808_pnld_ambient_rates.csv")

# selecting only the columns that I need
dat_a <- dat_a %>%
  select(chamber, total_DW_g, total_AFDW_g, rock_SA_m2:nutrient_trt)

dat_e <- read_csv("data/210808_pineland_a_e/210808_pnld_enriched_rates.csv")

# selecting only the columns that I need
dat_e <- dat_e %>%
  select(chamber, total_DW_g, total_AFDW_g, rock_SA_m2:nutrient_trt)


# binding the rows of the two dataframes together
dat <- bind_rows(dat_a, dat_e)
# adding in nutrient enriched dummy variable column (0 = ambient, 1 = enriched nutrients)
dat <- dat %>%
  mutate(nutrients = ifelse(nutrient_trt == "ambient", 0, 1))

# making ER values positive for subsequent log-transformations
dat <- dat %>%
  mutate(ER_mg_d = -1 * ER_mg_d, ER_mgO2_d_gAFDW = -1 * ER_mgO2_d_gAFDW, ER_mgO2_d_m2 = -1 * ER_mgO2_d_m2)

write_csv(dat, "data/main_analysis/210808_pnld_A_E_rates.csv")

#View(dat)

```

# AFDW-normalized Metabolic rates
### GPP vs temperature and nutrients
```{r}
# linear model of GPP and temp
GPP_AFDW_mod <- lm(log(GPP_mgO2_d_gAFDW) ~ temp_C*nutrients, dat)
summary(GPP_AFDW_mod)

# checking that model assumptions are met
plot(GPP_AFDW_mod)
# the scale-location plot may indicate that there is a potential issue with homoscedasticity. Q-Q plot also indicates possible heavy tails in the distribution.

# interpreting effect size of temperature on GPP
exp(coef(GPP_AFDW_mod)["temp_C"])
# looks that GPP increases about 4% for every 1C temp increase

GPP_AFDW_plot <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

GPP_AFDW_plot

ggsave("plots/210808_pineland_ambient_enriched/210808_GPP_AFDW_plot.png", plot = GPP_AFDW_plot, width = 7, height = 3.5)

# plotting an over-fitted line just to see what the resulting plot is. Looks very messy, and there is no clear trend.
GPP_AFDW_smooth_plot <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth() +
  stat_regline_equation() +
  scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

GPP_AFDW_smooth_plot
```
Temperature has signficant effect on GPP, with 1C temperature increases corresponding to a 4% increase in GPP.

### ER vs temperature and nutrients
```{r}
ER_AFDW_mod <- lm(log(ER_mgO2_d_gAFDW) ~ temp_C*nutrients, dat)
summary(ER_AFDW_mod)

# checking that model assumptions are met
plot(ER_AFDW_mod)
# Q-q plot shows that there are potentially heavy tails in the distribution.

# interpreting effect size of temperature on ER
exp(coef(ER_AFDW_mod)["temp_C"])
# ER increases by 9% for every 1C temp increase.

dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~gAFDW^{-1})))


# simplified plot w/o axis labels for use in plotly--This is used to identify which chambers correspond to outlier datapoints
p <- dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(title = "ER response to Temperature")

ER_plot <- plotly::ggplotly(p)
ER_plot

```
Temp was a significant predictor of AFDW-normalized ER, with a 1C temp increase corresponding to a 9% increase in ER.

### NEP vs. temp and nutrients
```{r}
# linear model of NEP and temp
NEP_AFDW_mod <- lm(log(NEP_mgO2_d_gAFDW) ~ temp_C*nutrients, dat)
summary(NEP_AFDW_mod)

dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~gAFDW)))

# simplified plot w/o axis labels for use in plotly--This is used to identify which chambers correspond to outlier datapoints
p <- dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(title = "NEP response to Temperature")

NEP_plot <- plotly::ggplotly(p)
NEP_plot

```

Chambers 14 an 6 seem like potential outliers for GPP, ER, and NEP. Chamber 14 highest dry weight and AFDW of all the chambers, perhaps these extreme values are affecting the AFDW-normalized rates. Chamber 6 has the second lowest AFDW, and the largest rock by volume, so this may potentially be affecting metabolic rates.

It also appears that temperature had the only significant effect (increasing both GPP and ER but not NEP). There are some potential issues with violations of linear model assumptions--particularly with normality--that need to be addressed.


# SA-normalized metabolic rates
### GPP vs temperature and nutrients model
```{r}
# linear model of GPP and temp
GPP_SA_mod <- lm(log(GPP_mgO2_d_m2) ~ temp_C*nutrients, dat)
summary(GPP_SA_mod)

GPP_SA_plot <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "GPP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~m^{-2})))
GPP_SA_plot

# making simplified plot w/o expressions for use in plotly function
GPP_SA <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  theme_classic()
# interactive plotly plot to get info on individual data points
GPP_SA_plotly <- plotly::ggplotly(GPP_SA)
GPP_SA_plotly
```

### ER vs temp and nutrients
```{r}
# linear model of GPP and temp
ER_SA_mod <- lm(log(ER_mgO2_d_m2) ~ temp_C*nutrients, dat)
summary(ER_SA_mod)

# checking that model assumptions are met
plot(ER_SA_mod)

# interpreting effect size of temperature on ER
exp(coef(ER_SA_mod)["temp_C"])
# ER increases by 4% for every 1C temp increase.

dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_m2, color = nutrient_trt)) +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~m^{-2})))

# making simplified plot without expressions in the axis labels for plotly
ER_SA_plot <- dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_m2, color = nutrient_trt)) +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic()

# ploting interactive plot in plotly to identify which datapoints correspond to which chambers
ER_SA_plotly <- plotly::ggplotly(ER_SA_plot)
ER_SA_plotly

```

### NEP vs temp and nutrients
```{r}
# linear model of NEP and temp
NEP_SA_mod <- lm(log(NEP_mgO2_d_m2) ~ temp_C*nutrients, dat)
summary(NEP_SA_mod)

# checking that model assumptions are met
plot(NEP_SA_mod)

dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~m^{-2})))

# making simplified plot without expressions in the axis labels for plotly
NEP_SA_plot <- dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  theme_classic()

# ploting interactive plot in plotly to identify which datapoints correspond to which chambers
NEP_SA_plotly <- plotly::ggplotly(NEP_SA_plot)
NEP_SA_plotly
```


# Plotting additional Variables to asses outlier datapoints
```{r}
#View(dat)
# plot of total AFDW of all chambers
dat %>%
  ggplot(aes(x = chamber, y = total_AFDW_g)) +
  geom_col() +
  theme_classic()

# plot of total DW of all chambers
dat %>%
  ggplot(aes(x = chamber, y = total_DW_g)) +
  geom_col() +
  theme_classic()

# plot of rock surface area of all chambers
dat %>%
  ggplot(aes(x = chamber, y = rock_SA_m2)) +
  geom_col() +
  theme_classic()

# calculating average total AFDW for thi incubation to compare to that of other incubations
mean(dat$total_AFDW_g)


### calculating average AFDW/m2 for the ambient and enriched incubations
avg_AFDW <- dat %>%
  filter(nutrient_trt == "ambient") %>% # subsetting to only include ambient since same rocks were used for ambient and enriched and would artificially double the sample size used to calculate the sd
  summarize(avg_AFDW_g_m2 = mean(total_AFDW_g/rock_SA_m2), sd_AFDW_g_m2 = sd(total_AFDW_g/rock_SA_m2))
avg_AFDW

```

# Double Logarithmic graphs (as specified in section 7.3.3 of del Giorgio & Williams 2005)

## GPP
```{r}
# linear model of log-transformed GPP (biomass-normalized) and log-transformed temperature

log_GPP_AFDW_mod <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat)
summary(log_GPP_AFDW_mod)

plot(log_GPP_AFDW_mod)
###### Model interpretations ######

# interpreting effect size of temperature effect
exp(coef(log_GPP_AFDW_mod)["log(temp_C)"])


GPP_log_log <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

ggsave("plots/210808_pineland_ambient_enriched/210808_GPP_log_log_plot.png", plot = GPP_log_log, width = 4, height = 2.5)
```

## ER
```{r}
# linear model of log-transformed ER rates (biomass normalized) and log-transformed temperature
log_ER_AFDW_mod <- lm(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat)
summary(log_ER_AFDW_mod)

plot(log_ER_AFDW_mod)

###### Model interpretations ######

# interpreting effect size of temperature effect
exp(coef(log_ER_AFDW_mod)["log(temp_C)"])


ER_log_log <- dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~gAFDW^{-1})))

ggsave("plots/210808_pineland_ambient_enriched/210808_ER_log_log_plot.png", plot = ER_log_log, width = 4, height = 2.5)
```

## NEP
```{r}
# linear model of log-transformed NEP (biomass normalized) and log-transformed temp
log_NEP_AFDW_mod <- lm(log(NEP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat)
summary(log_NEP_AFDW_mod)

plot(log_NEP_AFDW_mod)

# no significant predictors in the model


NEP_log_log <- dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~gAFDW)))

ggsave("plots/210808_pineland_ambient_enriched/210808_NEP_log_log_plot.png", plot = NEP_log_log, width = 4, height = 2.5)
```

