---
title: "211010 Pineland A & E Analysis"
author: "Nick Framsted"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# Importing and Processing data
```{r, include = FALSE}
# importing and cleaning up metabolism data from both ambient and enriched incubations
dat_a <- read_csv("data/211010_pineland_a_e/211010_pnld_ambient_rates.csv")

# selecting only the columns that I need
dat_a <- dat_a %>%
  select(chamber, total_DW_g, total_AFDW_g, rock_SA_m2:nutrient_trt)

dat_e <- read_csv("data/211010_pineland_a_e/211010_pnld_enriched_rates.csv")

# selecting only the columns that I need
dat_e <- dat_e %>%
  select(chamber, total_DW_g, total_AFDW_g, rock_SA_m2:nutrient_trt)


# binding the rows of the two dataframes together
dat <- bind_rows(dat_a, dat_e)
# adding in nutrient enriched dummy variable column (0 = ambient, 1 = enriched nutrients)
dat <- dat %>%
  mutate(nutrients = ifelse(nutrient_trt == "ambient", 0, 1))

# making ER values positive for subsequent log-transformations
dat <- dat %>%
  mutate(ER_mg_d = -1 * ER_mg_d, ER_mgO2_d_gAFDW = -1 * ER_mgO2_d_gAFDW, ER_mgO2_d_m2 = -1 * ER_mgO2_d_m2)

write_csv(dat, "data/main_analysis/211010_pnld_A_E_rates.csv")

#View(dat)

```

# AFDW-normalized Metabolic rates
### GPP vs temperature and nutrients
```{r}
# linear model of GPP and temp
GPP_AFDW_mod <- lm(log(GPP_mgO2_d_gAFDW) ~ temp_C*nutrients, dat)
summary(GPP_AFDW_mod)

# checking that model assumptions are met
plot(GPP_AFDW_mod)
# the scale-location plot may indicate that there is a potential issue with homoscedasticity. Q-Q plot also indicates possible heavy tails in the distribution.

# interpreting effect size of temperature on GPP
exp(coef(GPP_AFDW_mod)["temp_C"])
# looks that GPP increases about 4% for every 1C temp increase

GPP_AFDW_plot <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

GPP_AFDW_plot

ggsave("plots/211010_pineland_ambient_enriched/211010_GPP_AFDW_plot.png", plot = GPP_AFDW_plot, width = 7, height = 3.5)


```

### ER vs temperature and nutrients
```{r}
ER_AFDW_mod <- lm(log(ER_mgO2_d_gAFDW) ~ temp_C*nutrients, dat)
summary(ER_AFDW_mod)

# checking that model assumptions are met
plot(ER_AFDW_mod)
# Q-q plot shows that there are potentially heavy tails in the distribution.

# interpreting effect size of temperature on ER
exp(coef(ER_AFDW_mod)["temp_C"])
# ER increases by 9% for every 1C temp increase.

dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~gAFDW^{-1})))


```

### NEP vs. temp and nutrients
```{r}
# linear model of NEP and temp
NEP_AFDW_mod <- lm(log(NEP_mgO2_d_gAFDW) ~ temp_C*nutrients, dat)
summary(NEP_AFDW_mod)

dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~gAFDW)))


```

# SA-normalized metabolic rates
### GPP vs temperature and nutrients model
```{r}
# linear model of GPP and temp
GPP_SA_mod <- lm(log(GPP_mgO2_d_m2) ~ temp_C*nutrients, dat)
summary(GPP_SA_mod)

GPP_SA_plot <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "GPP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~m^{-2})))
GPP_SA_plot


```

### ER vs temp and nutrients
```{r}
# linear model of GPP and temp
ER_SA_mod <- lm(log(ER_mgO2_d_m2) ~ temp_C*nutrients, dat)
summary(ER_SA_mod)

# checking that model assumptions are met
plot(ER_SA_mod)

# interpreting effect size of temperature on ER
exp(coef(ER_SA_mod)["temp_C"])
# ER increases by 4% for every 1C temp increase.

dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~m^{-2})))

```

### NEP vs temp and nutrients
```{r}
# linear model of NEP and temp
NEP_SA_mod <- lm(log(NEP_mgO2_d_m2) ~ temp_C*nutrients, dat)
summary(NEP_SA_mod)

# checking that model assumptions are met
plot(NEP_SA_mod)

dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_m2, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~m^{-2})))

```

# Plotting additional Variables to asses outlier datapoints
```{r}
#View(dat)
# plot of total AFDW of all chambers
dat %>%
  ggplot(aes(x = chamber, y = total_AFDW_g)) +
  geom_col() +
  theme_classic()

# plot of total DW of all chambers
dat %>%
  ggplot(aes(x = chamber, y = total_DW_g)) +
  geom_col() +
  theme_classic()

# plot of rock surface area of all chambers
dat %>%
  ggplot(aes(x = chamber, y = rock_SA_m2)) +
  geom_col() +
  theme_classic()

# calculating average total AFDW for thi incubation to compare to that of other incubations
mean(dat$total_AFDW_g)


### calculating average AFDW/m2 for the ambient and enriched incubations
avg_AFDW <- dat %>%
  filter(nutrient_trt == "ambient") %>% # subsetting to only include ambient since same rocks were used for ambient and enriched and would artificially double the sample size used to calculate the sd
  summarize(avg_AFDW_g_m2 = mean(total_AFDW_g/rock_SA_m2), sd_AFDW_g_m2 = sd(total_AFDW_g/rock_SA_m2))
avg_AFDW

```

# Double Logarithmic graphs (as specified in section 7.3.3 of del Giorgio & Williams 2005)

## GPP
```{r}
# linear model of log-transformed GPP (biomass-normalized) and log-transformed temperature

log_GPP_AFDW_mod <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat)
summary(log_GPP_AFDW_mod)

plot(log_GPP_AFDW_mod)
###### Model interpretations ######

# interpreting effect size of temperature effect
exp(coef(log_GPP_AFDW_mod)["log(temp_C)"])


GPP_log_log <- dat %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

ggsave("plots/211010_pineland_ambient_enriched/211010_GPP_log_log_plot.png", plot = GPP_log_log, width = 4, height = 2.5)
```

## ER
```{r}
# linear model of log-transformed ER rates (biomass normalized) and log-transformed temperature
log_ER_AFDW_mod <- lm(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat)
summary(log_ER_AFDW_mod)

plot(log_ER_AFDW_mod)

###### Model interpretations ######

# interpreting effect size of temperature effect
exp(coef(log_ER_AFDW_mod)["log(temp_C)"])


ER_log_log <- dat %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #stat_regline_equation() +
  theme_classic() +
  labs(title = "ER response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~gAFDW^{-1})))

ggsave("plots/211010_pineland_ambient_enriched/211010_ER_log_log_plot.png", plot = ER_log_log, width = 4, height = 2.5)
```

## NEP
```{r}
# linear model of log-transformed NEP (biomass normalized) and log-transformed temp
log_NEP_AFDW_mod <- lm(log(NEP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, dat)
summary(log_NEP_AFDW_mod)

plot(log_NEP_AFDW_mod)

# no significant predictors in the model


NEP_log_log <- dat %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, color = nutrient_trt)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #stat_regline_equation() +
  theme_classic() +
  labs(title = "NEP response to Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~gAFDW)))

ggsave("plots/211010_pineland_ambient_enriched/211010_NEP_log_log_plot.png", plot = NEP_log_log, width = 4, height = 2.5)
```