---
title: "211128_pnld_ambient"
author: "Nick Framsted"
date: "12/27/2021"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# Importing and Processing data
```{r, include = FALSE}
dat <- read_csv("data/211128_pineland_a_e/211128_pineland_ambient_O2.csv", skip = 1)
#head(dat)
#View(dat)

# removing bottom row of NAs from dataframe
dat <- dat[-48493,]
#View(dat)

# Creating tank_channel column to assign chambers to individual data points in DO data by concatenating Device Serial and Channel columns. Keeping original columns as well. 
dat <- dat %>%
 unite(tank_channel, Device_Serial, Channel, sep = "_", remove = FALSE)


#View(dat)
#head(dat)



# filtering out rows with error codes, and adding in chamber numbers based the tank and channel number cross referenced against incubation notes
## double checked for correctness on 12/28/21 NTF

dat_2 <- dat %>%
  filter(Error == 0 & !is.na(Value)) %>%
  mutate(
    chamber = case_when(
    tank_channel == "Tank 1_1"  ~ "11",
    tank_channel == "Tank 1_2"  ~ "13",
    tank_channel == "Tank 1_3" ~ "15",
    tank_channel == "Tank 1_4" ~ "16",
    tank_channel == "Tank 2_1" ~ "1",
    tank_channel == "Tank 2_2" ~ "3",
    tank_channel == "Tank 2_3" ~ "4",
    tank_channel == "Tank 2_4" ~ "6",
    tank_channel == "Tank 3_1" ~ "5",
    tank_channel == "Tank 3_2" ~ "8",
    tank_channel == "Tank 3_3" ~ "12",
    tank_channel == "Tank 3_4" ~ "14",
    tank_channel == "Tank 4_1" ~ "2",
    tank_channel == "Tank 4_2" ~ "7",
    tank_channel == "Tank 4_3" ~ "9",
    tank_channel == "Tank 4_4" ~ "10",
  )) %>%
   mutate(treatment = case_when(
     Device_Serial == "Tank 1" ~ "12.5C",
     Device_Serial == "Tank 2" ~ "9.5C",
     Device_Serial == "Tank 3" ~ "15.5C",
     Device_Serial == "Tank 4" ~ "18.5C"))
 
#View(dat_2)
head(dat_2)

# renaming oxygen data column and the delta_T columns to include units
dat_2 <- dat_2 %>%
  rename(oxygen = Value, time = Time, date = Date, temp = Temp, delta_T_min = delta_t)
head(dat_2)

### Merging DO data with chamber data ###

# selecting only columns I need, and changing chamber column from a factor to a double
dat_subset <- dat_2 %>%
  dplyr::select(date, time, delta_T_min, oxygen, temp, chamber, treatment) %>%
  filter(delta_T_min > 900) #filtering out NAs and also test DO data from the day before
#View(dat_subset)
  dat_subset$chamber <- as.numeric(dat_subset$chamber)

# importing in other complementary data associated with chambers
dat_chambers <- read_csv("data/211128_pineland_a_e/211128_pineland_A_E_datasheet.csv")
#str(dat_chambers)
dat_chambers$chamber <- as.numeric(dat_chambers$chamber)

# removing empty rows
dat_chambers <- dat_chambers %>%
  filter(!is.na(chamber)) #filtering out NAs
#View(dat_chambers)

# changing remeasured AFDW value for chamber 13 back to the original since this value was negative and was not realistic.
dat_chambers <- dat_chambers %>%
  mutate(total_AFDW_final = ifelse(total_AFDW_g_2 < 0, 0.392,total_AFDW_g_2))
#View(dat_chambers)

# joining DO data and chamber data by "chamber" column
dat_full <- full_join(dat_subset, dat_chambers, by = "chamber")
#View(dat_full)

# converting oxygen concentration (mg/L) to oxygen mass (mg) by multipying by water volumes in chambers
dat_full <- dat_full %>%
  mutate(oxygen_mass_mg = oxygen * water_volume_L) %>%
  filter(!is.na(oxygen_mass_mg)) # filtering out rows with NAs for DO data
#View(dat_full)
```

# Calculating % water in AFDW samples
```{r}
# percent water content of first round of AFDW samples that were ashed at the incorrect temperature. This was used to generate Dry wts for a second round of samples ashed at the correct temperature, but were not weighed for their dry wts before ashing.
dat_chambers %>%
  mutate(AFDW_perc_water = (SWW_g - SDW_60C_g)/SWW_g) %>%
  summarise(avg_perc_water = mean(AFDW_perc_water), sd_perc_water = sd(AFDW_perc_water))
```


# Calculating % difference between ashing temps
```{r}
# The UCD muffle furnace in PES building was reporting the incorrect temperature and thus samples were ashed at incorrect temps. These samples were ashed at 270C instead of 500C and then were subsequently ashed at 580C. Below are the calculated percent differences in AFDW between the two ashing temps:

dat_chambers %>%
  mutate(SCW_diff_270_500 = SCW_270C_g - SCW_500C_g_2, SCW_diff_580_500 = SCW_580C_g - SCW_500C_g_2) %>%
  summarise(avg_diff_270 = mean(SCW_diff_270_500), sd_diff_270 = sd(SCW_diff_270_500), avg_diff_580C = mean(SCW_diff_580_500), sd_diff_580C = sd(SCW_diff_580_500))
```

# Calculating % change in AFDW between different ashing temps
```{r}
dat_chambers %>%
  mutate(AFDW_270 = (total_wet_wt_g/SWW_g) * (SDW_60C_g - SCW_270C_g), AFDW_500 = (total_wet_wt_g/SWW_g_2) * (SDW_60C_g_2 - SCW_500C_g_2), AFDW_580 = (total_wet_wt_g/SWW_g) * (SDW_60C_g - SCW_580C_g)) %>%
  filter(AFDW_500 > 0 & chamber != 2) %>% # filtering out outlying AFDW data points (chamber 2 and 13)
  mutate(perc_diff_270 = (AFDW_500-AFDW_270)/AFDW_500, perc_diff_580 = (AFDW_500-AFDW_580)/AFDW_500) %>%
  summarise(avg_perc_diff_270 = mean(perc_diff_270), sd_perc_diff_270 = sd(perc_diff_270), avg_perc_diff_580 = mean(perc_diff_580), sd_perc_diff_580 = sd(perc_diff_580))
```


# Plotting temp and DO in chambers
```{r echo = FALSE}
# making treatment a factor variable and re-ordering it for plots
dat_full$treatment <- factor(dat_full$treatment, levels = c("9.5C", "12.5C", "15.5C", "18.5C"))
# making chamber a factor so that it's a discrete value for coloring datapoints in plot
dat_full$chamber <- as.factor(dat_full$chamber)

# plotting Do concentrations in each chamber
DO_conc_plot <- dat_full %>%
  ggplot(aes(x = delta_T_min, y = oxygen, color = chamber)) +
  geom_line(size = 1.5) +
  scale_color_viridis_d() +
  facet_wrap(~treatment) +
  labs(y = expression(Oxygen~Concentration~(mg~L^{-1})), x = "Elapsed Time (min)", title = "Pineland Ambient Incubation DO") +
  theme_bw(base_size = 20)

DO_conc_plot

ggsave("plots/211128_pineland_ambient_enriched/211128_pineland_ambient_DO_conc_plot.png", plot = DO_conc_plot, width = 10, height = 7)

# plotting temperature for each chamber
temp_plot <- dat_full %>%
  ggplot(aes(x = delta_T_min, y = temp, color = chamber)) +
  geom_line(size = 1.5) +
  scale_color_viridis_d() +
  facet_wrap(~treatment) +
  labs(y = expression('Temperature (\u00B0C)'), x = "Elapsed Time (min)", title = "Pineland Ambient Incubation Temperature") +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")

temp_plot

ggsave("plots/211128_pineland_ambient_enriched/211128_pineland_ambient_temp_plot.png", plot = temp_plot, width = 10, height = 7)

# plotting DO mass (mg) for each chamber
DO_mass_plot <- dat_full %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg, color = chamber)) +
  geom_line(size = 1.5) +
  scale_color_viridis_d() +
  facet_wrap(~treatment) +
  labs(y = expression(Oxygen~Mass~(mg)), x = "Elapsed Time (min)", title = "Pineland Ambient Incubation DO Mass") +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")

  DO_mass_plot
  
ggsave("plots/211128_pineland_ambient_enriched/211128_pineland_ambient_DO_mass_plot.png", plot = DO_mass_plot, width = 10, height = 7)

# making interactive plot with plotly to help identity data points for rate calculations below
p<- plotly::ggplotly(DO_mass_plot)
p
# exporting interactive plot to html
htmlwidgets::saveWidget(as_widget(p), "plots/211128_pineland_ambient_enriched/211128_pineland_ambient_DO_mass_plot.html")

```

## Metabolic Rate Calculations
### Tank 2; 9.5C
#### Chamber 1
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_1 <- dat_full %>%
  filter(chamber == "1" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_1_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_1)
summary(pnld_a_1_NEP_mod)
pnld_a_1_NEP_mod$coefficients[2]
pnld_a_1_NEP_slope <- unname(pnld_a_1_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_1_NEP <- pnld_a_1_NEP_slope * 1440
pnld_a_1_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_1 <- dat_full %>%
  filter(chamber =="1" & delta_T_min >= 1013)

pnld_a_1_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_1)
summary(pnld_a_1_ER_mod)
pnld_a_1_ER_mod$coefficients[2]
pnld_a_1_ER_slope <- unname(pnld_a_1_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_1_ER <- pnld_a_1_ER_slope * 1440
pnld_a_1_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_1_plot <- dat_full %>%
  filter(chamber == "1") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_1_NEP_mod$coefficients[1], slope = pnld_a_1_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_1_ER_mod$coefficients[1], slope = pnld_a_1_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_1_plot
```
##### has R2 below 0.9

#### Chamber 3
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_3 <- dat_full %>%
  filter(chamber == "3" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_3_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_3)
summary(pnld_a_3_NEP_mod)
pnld_a_3_NEP_mod$coefficients[2]
pnld_a_3_NEP_slope <- unname(pnld_a_3_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_3_NEP <- pnld_a_3_NEP_slope * 1440
pnld_a_3_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_3 <- dat_full %>%
  filter(chamber =="3" & delta_T_min >= 1013)

pnld_a_3_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_3)
summary(pnld_a_3_ER_mod)
pnld_a_3_ER_mod$coefficients[2]
pnld_a_3_ER_slope <- unname(pnld_a_3_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_3_ER <- pnld_a_3_ER_slope * 1440
pnld_a_3_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_3_plot <- dat_full %>%
  filter(chamber == "3") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_3_NEP_mod$coefficients[1], slope = pnld_a_3_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_3_ER_mod$coefficients[1], slope = pnld_a_3_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_3_plot
```

#### Chamber 4
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_4 <- dat_full %>%
  filter(chamber == "4" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_4_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_4)
summary(pnld_a_4_NEP_mod)
pnld_a_4_NEP_mod$coefficients[2]
pnld_a_4_NEP_slope <- unname(pnld_a_4_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_4_NEP <- pnld_a_4_NEP_slope * 1440
pnld_a_4_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_4 <- dat_full %>%
  filter(chamber =="4" & delta_T_min >= 1013)

pnld_a_4_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_4)
summary(pnld_a_4_ER_mod)
pnld_a_4_ER_mod$coefficients[2]
pnld_a_4_ER_slope <- unname(pnld_a_4_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_4_ER <- pnld_a_4_ER_slope * 1440
pnld_a_4_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_4_plot <- dat_full %>%
  filter(chamber == "4") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_4_NEP_mod$coefficients[1], slope = pnld_a_4_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_4_ER_mod$coefficients[1], slope = pnld_a_4_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_4_plot
```
##### has R2 near 0.9 (0.93 exactly)

#### Chamber 6
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_6 <- dat_full %>%
  filter(chamber == "6" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_6_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_6)
summary(pnld_a_6_NEP_mod)
pnld_a_6_NEP_mod$coefficients[2]
pnld_a_6_NEP_slope <- unname(pnld_a_6_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_6_NEP <- pnld_a_6_NEP_slope * 1440
pnld_a_6_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_6 <- dat_full %>%
  filter(chamber =="6" & delta_T_min >= 1013)

pnld_a_6_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_6)
summary(pnld_a_6_ER_mod)
pnld_a_6_ER_mod$coefficients[2]
pnld_a_6_ER_slope <- unname(pnld_a_6_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_6_ER <- pnld_a_6_ER_slope * 1440
pnld_a_6_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_6_plot <- dat_full %>%
  filter(chamber == "6") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_6_NEP_mod$coefficients[1], slope = pnld_a_6_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_6_ER_mod$coefficients[1], slope = pnld_a_6_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_6_plot
```

### Tank 1; 12.5C
#### Chamber 11
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_11 <- dat_full %>%
  filter(chamber == "11" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_11_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_11)
summary(pnld_a_11_NEP_mod)
pnld_a_11_NEP_mod$coefficients[2]
pnld_a_11_NEP_slope <- unname(pnld_a_11_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_11_NEP <- pnld_a_11_NEP_slope * 1440
pnld_a_11_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_11 <- dat_full %>%
  filter(chamber =="11" & delta_T_min >= 1013)

pnld_a_11_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_11)
summary(pnld_a_11_ER_mod)
pnld_a_11_ER_mod$coefficients[2]
pnld_a_11_ER_slope <- unname(pnld_a_11_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_11_ER <- pnld_a_11_ER_slope * 1440
pnld_a_11_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_11_plot <- dat_full %>%
  filter(chamber == "11") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_11_NEP_mod$coefficients[1], slope = pnld_a_11_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_11_ER_mod$coefficients[1], slope = pnld_a_11_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_11_plot
```

#### Chamber 13
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_13 <- dat_full %>%
  filter(chamber == "13" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_13_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_13)
summary(pnld_a_13_NEP_mod)
pnld_a_13_NEP_mod$coefficients[2]
pnld_a_13_NEP_slope <- unname(pnld_a_13_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_13_NEP <- pnld_a_13_NEP_slope * 1440
pnld_a_13_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_13 <- dat_full %>%
  filter(chamber =="13" & delta_T_min >= 1013)

pnld_a_13_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_13)
summary(pnld_a_13_ER_mod)
pnld_a_13_ER_mod$coefficients[2]
pnld_a_13_ER_slope <- unname(pnld_a_13_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_13_ER <- pnld_a_13_ER_slope * 1440
pnld_a_13_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_13_plot <- dat_full %>%
  filter(chamber == "13") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_13_NEP_mod$coefficients[1], slope = pnld_a_13_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_13_ER_mod$coefficients[1], slope = pnld_a_13_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_13_plot
```

#### Chamber 15
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_15 <- dat_full %>%
  filter(chamber == "15" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_15_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_15)
summary(pnld_a_15_NEP_mod)
pnld_a_15_NEP_mod$coefficients[2]
pnld_a_15_NEP_slope <- unname(pnld_a_15_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_15_NEP <- pnld_a_15_NEP_slope * 1440
pnld_a_15_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_15 <- dat_full %>%
  filter(chamber =="15" & delta_T_min >= 1013)

pnld_a_15_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_15)
summary(pnld_a_15_ER_mod)
pnld_a_15_ER_mod$coefficients[2]
pnld_a_15_ER_slope <- unname(pnld_a_15_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_15_ER <- pnld_a_15_ER_slope * 1440
pnld_a_15_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_15_plot <- dat_full %>%
  filter(chamber == "15") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_15_NEP_mod$coefficients[1], slope = pnld_a_15_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_15_ER_mod$coefficients[1], slope = pnld_a_15_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_15_plot
```

#### Chamber 16
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_16 <- dat_full %>%
  filter(chamber == "16" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_16_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_16)
summary(pnld_a_16_NEP_mod)
pnld_a_16_NEP_mod$coefficients[2]
pnld_a_16_NEP_slope <- unname(pnld_a_16_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_16_NEP <- pnld_a_16_NEP_slope * 1440
pnld_a_16_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_16 <- dat_full %>%
  filter(chamber =="16" & delta_T_min >= 1013)

pnld_a_16_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_16)
summary(pnld_a_16_ER_mod)
pnld_a_16_ER_mod$coefficients[2]
pnld_a_16_ER_slope <- unname(pnld_a_16_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_16_ER <- pnld_a_16_ER_slope * 1440
pnld_a_16_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_16_plot <- dat_full %>%
  filter(chamber == "16") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_16_NEP_mod$coefficients[1], slope = pnld_a_16_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_16_ER_mod$coefficients[1], slope = pnld_a_16_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_16_plot
```

### Tank 3; 15.5C
#### Chamber 5
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_5 <- dat_full %>%
  filter(chamber == "5" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_5_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_5)
summary(pnld_a_5_NEP_mod)
pnld_a_5_NEP_mod$coefficients[2]
pnld_a_5_NEP_slope <- unname(pnld_a_5_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_5_NEP <- pnld_a_5_NEP_slope * 1440
pnld_a_5_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_5 <- dat_full %>%
  filter(chamber =="5" & delta_T_min >= 1013)

pnld_a_5_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_5)
summary(pnld_a_5_ER_mod)
pnld_a_5_ER_mod$coefficients[2]
pnld_a_5_ER_slope <- unname(pnld_a_5_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_5_ER <- pnld_a_5_ER_slope * 1440
pnld_a_5_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_5_plot <- dat_full %>%
  filter(chamber == "5") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_5_NEP_mod$coefficients[1], slope = pnld_a_5_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_5_ER_mod$coefficients[1], slope = pnld_a_5_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_5_plot
```

#### Chamber 8
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_8 <- dat_full %>%
  filter(chamber == "8" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_8_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_8)
summary(pnld_a_8_NEP_mod)
pnld_a_8_NEP_mod$coefficients[2]
pnld_a_8_NEP_slope <- unname(pnld_a_8_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_8_NEP <- pnld_a_8_NEP_slope * 1440
pnld_a_8_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_8 <- dat_full %>%
  filter(chamber =="8" & delta_T_min >= 1013)

pnld_a_8_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_8)
summary(pnld_a_8_ER_mod)
pnld_a_8_ER_mod$coefficients[2]
pnld_a_8_ER_slope <- unname(pnld_a_8_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_8_ER <- pnld_a_8_ER_slope * 1440
pnld_a_8_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_8_plot <- dat_full %>%
  filter(chamber == "8") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_8_NEP_mod$coefficients[1], slope = pnld_a_8_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_8_ER_mod$coefficients[1], slope = pnld_a_8_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_8_plot
```

#### Chamber 12
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_12 <- dat_full %>%
  filter(chamber == "12" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_12_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_12)
summary(pnld_a_12_NEP_mod)
pnld_a_12_NEP_mod$coefficients[2]
pnld_a_12_NEP_slope <- unname(pnld_a_12_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_12_NEP <- pnld_a_12_NEP_slope * 1440
pnld_a_12_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_12 <- dat_full %>%
  filter(chamber =="12" & delta_T_min >= 1013)

pnld_a_12_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_12)
summary(pnld_a_12_ER_mod)
pnld_a_12_ER_mod$coefficients[2]
pnld_a_12_ER_slope <- unname(pnld_a_12_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_12_ER <- pnld_a_12_ER_slope * 1440
pnld_a_12_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_12_plot <- dat_full %>%
  filter(chamber == "12") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_12_NEP_mod$coefficients[1], slope = pnld_a_12_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_12_ER_mod$coefficients[1], slope = pnld_a_12_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_12_plot
```

#### Chamber 14
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_14 <- dat_full %>%
  filter(chamber == "14" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_14_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_14)
summary(pnld_a_14_NEP_mod)
pnld_a_14_NEP_mod$coefficients[2]
pnld_a_14_NEP_slope <- unname(pnld_a_14_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_14_NEP <- pnld_a_14_NEP_slope * 1440
pnld_a_14_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_14 <- dat_full %>%
  filter(chamber =="14" & delta_T_min >= 1013)

pnld_a_14_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_14)
summary(pnld_a_14_ER_mod)
pnld_a_14_ER_mod$coefficients[2]
pnld_a_14_ER_slope <- unname(pnld_a_14_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_14_ER <- pnld_a_14_ER_slope * 1440
pnld_a_14_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_14_plot <- dat_full %>%
  filter(chamber == "14") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_14_NEP_mod$coefficients[1], slope = pnld_a_14_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_14_ER_mod$coefficients[1], slope = pnld_a_14_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_14_plot
```

### Tank 4; 18.5C
#### Chamber 2
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_2 <- dat_full %>%
  filter(chamber == "2" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_2_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_2)
summary(pnld_a_2_NEP_mod)
pnld_a_2_NEP_mod$coefficients[2]
pnld_a_2_NEP_slope <- unname(pnld_a_2_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_2_NEP <- pnld_a_2_NEP_slope * 1440
pnld_a_2_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_2 <- dat_full %>%
  filter(chamber =="2" & delta_T_min >= 1013)

pnld_a_2_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_2)
summary(pnld_a_2_ER_mod)
pnld_a_2_ER_mod$coefficients[2]
pnld_a_2_ER_slope <- unname(pnld_a_2_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_2_ER <- pnld_a_2_ER_slope * 1440
pnld_a_2_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_2_plot <- dat_full %>%
  filter(chamber == "2") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_2_NEP_mod$coefficients[1], slope = pnld_a_2_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_2_ER_mod$coefficients[1], slope = pnld_a_2_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_2_plot
```

#### Chamber 7
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_7 <- dat_full %>%
  filter(chamber == "7" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_7_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_7)
summary(pnld_a_7_NEP_mod)
pnld_a_7_NEP_mod$coefficients[2]
pnld_a_7_NEP_slope <- unname(pnld_a_7_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_7_NEP <- pnld_a_7_NEP_slope * 1440
pnld_a_7_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_7 <- dat_full %>%
  filter(chamber =="7" & delta_T_min >= 1013)

pnld_a_7_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_7)
summary(pnld_a_7_ER_mod)
pnld_a_7_ER_mod$coefficients[2]
pnld_a_7_ER_slope <- unname(pnld_a_7_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_7_ER <- pnld_a_7_ER_slope * 1440
pnld_a_7_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_7_plot <- dat_full %>%
  filter(chamber == "7") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_7_NEP_mod$coefficients[1], slope = pnld_a_7_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_7_ER_mod$coefficients[1], slope = pnld_a_7_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_7_plot
```

#### Chamber 9
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_9 <- dat_full %>%
  filter(chamber == "9" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_9_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_9)
summary(pnld_a_9_NEP_mod)
pnld_a_9_NEP_mod$coefficients[2]
pnld_a_9_NEP_slope <- unname(pnld_a_9_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_9_NEP <- pnld_a_9_NEP_slope * 1440
pnld_a_9_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_9 <- dat_full %>%
  filter(chamber =="9" & delta_T_min >= 1013)

pnld_a_9_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_9)
summary(pnld_a_9_ER_mod)
pnld_a_9_ER_mod$coefficients[2]
pnld_a_9_ER_slope <- unname(pnld_a_9_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_9_ER <- pnld_a_9_ER_slope * 1440
pnld_a_9_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_9_plot <- dat_full %>%
  filter(chamber == "9") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_9_NEP_mod$coefficients[1], slope = pnld_a_9_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_9_ER_mod$coefficients[1], slope = pnld_a_9_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_9_plot
```

#### Chamber 10
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 933 to minute 1013 ##

pnld_a_10 <- dat_full %>%
  filter(chamber == "10" & delta_T_min >= 933 & delta_T_min <= 1013)

pnld_a_10_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_10)
summary(pnld_a_10_NEP_mod)
pnld_a_10_NEP_mod$coefficients[2]
pnld_a_10_NEP_slope <- unname(pnld_a_10_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_10_NEP <- pnld_a_10_NEP_slope * 1440
pnld_a_10_NEP


## now measuring ER from the period after the NEP portion of the incubation (oxygen curve stabilizes past minute 1013.
pnld_a_10 <- dat_full %>%
  filter(chamber =="10" & delta_T_min >= 1013)

pnld_a_10_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_a_10)
summary(pnld_a_10_ER_mod)
pnld_a_10_ER_mod$coefficients[2]
pnld_a_10_ER_slope <- unname(pnld_a_10_ER_mod$coefficients[2])



## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_a_10_ER <- pnld_a_10_ER_slope * 1440
pnld_a_10_ER


# plot of DO in this chamber to check that regressions match with data
pnld_a_10_plot <- dat_full %>%
  filter(chamber == "10") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_a_10_NEP_mod$coefficients[1], slope = pnld_a_10_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_a_10_ER_mod$coefficients[1], slope = pnld_a_10_ER_mod$coefficients[2])) +
  theme_classic()
pnld_a_10_plot
```

## Creating dataframe with metab rates
```{r}
# creating dataframe of metab rates to make barplot
pnld_a_rates <- data.frame(
  "NEP_mg_d" = c(pnld_a_1_NEP, pnld_a_2_NEP, pnld_a_3_NEP, pnld_a_4_NEP, pnld_a_5_NEP, pnld_a_6_NEP, pnld_a_7_NEP, pnld_a_8_NEP, pnld_a_9_NEP, pnld_a_10_NEP, pnld_a_11_NEP, pnld_a_12_NEP, pnld_a_13_NEP, pnld_a_14_NEP, pnld_a_15_NEP, pnld_a_16_NEP),
  "ER_mg_d" = c(pnld_a_1_ER, pnld_a_2_ER, pnld_a_3_ER, pnld_a_4_ER, pnld_a_5_ER, pnld_a_6_ER, pnld_a_7_ER, pnld_a_8_ER, pnld_a_9_ER, pnld_a_10_ER, pnld_a_11_ER, pnld_a_12_ER, pnld_a_13_ER, pnld_a_14_ER, pnld_a_15_ER,  pnld_a_16_ER),
  "chamber" = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16"))
head(pnld_a_rates)
#str(pnld_a_rates)

# formatting data types of columns
pnld_a_rates$NEP_mg_d <- as.numeric(pnld_a_rates$NEP_mg_d)
pnld_a_rates$ER_mg_d <- as.numeric(pnld_a_rates$ER_mg_d)

# calculating GPP from NEP and ER columns
pnld_a_rates <- pnld_a_rates %>%
  mutate(GPP_mg_d = NEP_mg_d - ER_mg_d)
head(pnld_a_rates)
#View(pnld_a_rates)
```

## Joining metab rates dataframe with chamber dataframe
```{r}

# changing chamber column in datasheet to character to be compatable to joining to pnld_a_rates
dat_chambers$chamber <- as.factor(dat_chambers$chamber)

# joining chamber and metabolic rate data
dat_final <- full_join(dat_chambers, pnld_a_rates, by = "chamber")
#View(dat_final)

# adding temperature treatment (discrete factor) and temp_C (continuous numeric) columns based off of tank column already in data
dat_final <- dat_final %>%
  mutate(
    treatment = case_when(
      Tank_a == "1" ~ "12.5C",
      Tank_a == "2" ~ "9.5C",
      Tank_a == "3" ~ "15.5C",
      Tank_a == "4" ~ "18.5C"
    )) %>%
  mutate(
    temp_C = case_when(
           Tank_a == "1" ~ 12.5,
           Tank_a == "2" ~ 9.5,
           Tank_a == "3" ~ 15.5,
           Tank_a == "4" ~ 18.5
           ))


#View(dat_final)
```

## Normalizing metabolic rates
```{r}
# normalizing metabolic rates by AFDW and by rock surface area
dat_final <- dat_final %>%
  mutate(GPP_mgO2_d_gAFDW = GPP_mg_d/total_AFDW_final, NEP_mgO2_d_gAFDW = NEP_mg_d/total_AFDW_final, ER_mgO2_d_gAFDW = ER_mg_d/total_AFDW_final) %>%
  mutate(GPP_mgO2_d_m2 = GPP_mg_d/rock_SA_m2, NEP_mgO2_d_m2 = NEP_mg_d/rock_SA_m2, ER_mgO2_d_m2 = ER_mg_d/rock_SA_m2)

#View(dat_final)
```

# Exporting metabolic rate data for further analysis
```{r}
# making identifier column for joining to ambient incubation data for nutrient effect analysis
dat_final <- dat_final %>%
  mutate(nutrient_trt = "ambient")
# exporting data for final analysis
write_csv(dat_final, "data/211128_pineland_a_e/211128_pnld_ambient_rates.csv")
```