---
title: "211128_pnld_enriched"
author: "Nick Framsted"
date: "12/29/2021"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# Importing and Processing data
```{r, include = FALSE}
dat <- read_csv("data/211128_pineland_a_e/211128_pineland_enriched_O2.csv", skip = 1)
#head(dat)
#View(dat)

# removing bottom row of NAs from dataframe
dat <- dat[-39061,]
#View(dat)

# Creating tank_channel column to assign chambers to individual data points in DO data by concatenating Device Serial and Channel columns. Keeping original columns as well. 
dat <- dat %>%
 unite(tank_channel, Device_Serial, Channel, sep = "_", remove = FALSE)


#View(dat)
#head(dat)



# filtering out rows with error codes, and adding in chamber numbers based the tank and channel number cross referenced against incubation notes
## double checked for correctness on 11/02/21 NTF

dat_2 <- dat %>%
  filter(Error == 0 & !is.na(Value)) %>%
  mutate(
    chamber = case_when(
    tank_channel == "Tank 1_1"  ~ "5",
    tank_channel == "Tank 1_2"  ~ "10",
    tank_channel == "Tank 1_3" ~ "12",
    tank_channel == "Tank 1_4" ~ "13",
    tank_channel == "Tank 2_1" ~ "3",
    tank_channel == "Tank 2_2" ~ "7",
    tank_channel == "Tank 2_3" ~ "15",
    tank_channel == "Tank 2_4" ~ "16",
    tank_channel == "Tank 3_1" ~ "1",
    tank_channel == "Tank 3_2" ~ "2",
    tank_channel == "Tank 3_3" ~ "4",
    tank_channel == "Tank 3_4" ~ "14",
    tank_channel == "Tank 4_1" ~ "6",
    tank_channel == "Tank 4_2" ~ "8",
    tank_channel == "Tank 4_3" ~ "9",
    tank_channel == "Tank 4_4" ~ "11",
  )) %>%
   mutate(treatment = case_when(
     Device_Serial == "Tank 1" ~ "12.5C",
     Device_Serial == "Tank 2" ~ "9.5C",
     Device_Serial == "Tank 3" ~ "15.5C",
     Device_Serial == "Tank 4" ~ "18.5C"))
 
#View(dat_2)
head(dat_2)

# renaming oxygen data column and the delta_T columns to include units
dat_2 <- dat_2 %>%
  rename(oxygen = Value, time = Time, date = Date, temp = Temp, delta_T_min = delta_t)
head(dat_2)

### Merging DO data with chamber data ###

# selecting only columns I need, and changing chamber column from a factor to a double
dat_subset <- dat_2 %>%
  select(date, time, delta_T_min, oxygen, temp, chamber, treatment) %>%
  filter(delta_T_min > 1000) #filtering out NAs and also test DO data from the day before
#View(dat_subset)
  dat_subset$chamber <- as.numeric(dat_subset$chamber)

# importing in other complementary data associated with chambers
dat_chambers <- read_csv("data/211128_pineland_a_e/211128_pineland_A_E_datasheet.csv")
#str(dat_chambers)
dat_chambers$chamber <- as.numeric(dat_chambers$chamber)

# removing empty rows
dat_chambers <- dat_chambers %>%
  filter(!is.na(chamber))
#View(dat_chambers)

# changing remeasured AFDW value for chamber 13 back to the original since this value was negative and was not realistic.
dat_chambers <- dat_chambers %>%
  mutate(total_AFDW_final = ifelse(total_AFDW_g_2 < 0, 0.392,total_AFDW_g_2))
#View(dat_chambers)

# joining DO data and chamber data by "chamber" column
dat_full <- full_join(dat_subset, dat_chambers, by = "chamber")
#View(dat_full)

# converting oxygen concentration (mg/L) to oxygen mass (mg) by multipying by water volumes in chambers
dat_full <- dat_full %>%
  mutate(oxygen_mass_mg = oxygen * water_volume_L) %>%
  filter(!is.na(oxygen_mass_mg)) # filtering out rows with NAs for DO data
#View(dat_full)
```

# Plotting temp and DO in chambers
```{r echo = FALSE}
# making treatment a factor variable and re-ordering it for plots
dat_full$treatment <- factor(dat_full$treatment, levels = c("9.5C", "12.5C", "15.5C", "18.5C"))
# making chamber a factor so that it's a discrete value for coloring datapoints in plot
dat_full$chamber <- as.factor(dat_full$chamber)

# plotting Do concentrations in each chamber
DO_conc_plot <- dat_full %>%
  ggplot(aes(x = delta_T_min, y = oxygen, color = chamber)) +
  geom_line(size = 1.5) +
  scale_color_viridis_d() +
  facet_wrap(~treatment) +
  labs(y = expression(Oxygen~Concentration~(mg~L^{-1})), x = "Elapsed Time (min)", title = "Pineland Enriched Incubation DO") +
  theme_bw(base_size = 20)

DO_conc_plot

ggsave("plots/211128_pineland_ambient_enriched/211128_pineland_enriched_DO_conc_plot.png", plot = DO_conc_plot, width = 10, height = 11)

# plotting temperature for each chamber
temp_plot <- dat_full %>%
  ggplot(aes(x = delta_T_min, y = temp, color = chamber)) +
  geom_line(size = 1.5) +
  scale_color_viridis_d() +
  facet_wrap(~treatment) +
  labs(y = expression('Temperature (\u00B0C)'), x = "Elapsed Time (min)", title = "Pineland enriched Incubation Temperature") +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")

temp_plot

ggsave("plots/211128_pineland_ambient_enriched/211128_pineland_enriched_temp_plot.png", plot = temp_plot, width = 10, height = 11)

# plotting DO mass (mg) for each chamber
DO_mass_plot <- dat_full %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg, color = chamber)) +
  geom_line(size = 1.5) +
  scale_color_viridis_d() +
  facet_wrap(~treatment) +
  labs(y = expression(Oxygen~Mass~(mg)), x = "Elapsed Time (min)", title = "Pineland enriched Incubation DO Mass") +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")

  DO_mass_plot
  
ggsave("plots/211128_pineland_ambient_enriched/211128_pineland_enriched_DO_mass_plot.png", plot = DO_mass_plot, width = 10, height = 11)

# making interactive plot with plotly to help identity data points for rate calculations below
p<- plotly::ggplotly(DO_mass_plot)
p
# exporting interactive plot to html
htmlwidgets::saveWidget(as_widget(p), "plots/211128_pineland_ambient_enriched/211128_pineland_enriched_DO_mass_plot.html")

```

## Metabolic Rate Calculations
### Tank 2; 9.5C
#### Chamber 5
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_5 <- dat_full %>%
  filter(chamber == "5" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_5_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_5)
summary(pnld_e_5_NEP_mod)
pnld_e_5_NEP_mod$coefficients[2]
pnld_e_5_NEP_slope <- unname(pnld_e_5_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_5_NEP <- pnld_e_5_NEP_slope * 1440
pnld_e_5_NEP

######################## ER rate calculations

pnld_e_5 <- dat_full %>%
  filter(chamber == "5" & delta_T_min >= 1206)

pnld_e_5_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_5)
summary(pnld_e_5_ER_mod)
pnld_e_5_ER_mod$coefficients[2]
pnld_e_5_ER_slope <- unname(pnld_e_5_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_5_ER <- pnld_e_5_ER_slope * 1440
pnld_e_5_ER

######################## GPP
pnld_e_5_GPP <- pnld_e_5_NEP - pnld_e_5_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_5_plot <- dat_full %>%
  filter(chamber == "5") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_5_NEP_mod$coefficients[1], slope = pnld_e_5_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_5_ER_mod$coefficients[1], slope = pnld_e_5_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_5_plot
```

#### Chamber 10
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_10 <- dat_full %>%
  filter(chamber == "10" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_10_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_10)
summary(pnld_e_10_NEP_mod)
pnld_e_10_NEP_mod$coefficients[2]
pnld_e_10_NEP_slope <- unname(pnld_e_10_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_10_NEP <- pnld_e_10_NEP_slope * 1440
pnld_e_10_NEP

######################## ER rate calculations

pnld_e_10 <- dat_full %>%
  filter(chamber == "10" & delta_T_min >= 1206)

pnld_e_10_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_10)
summary(pnld_e_10_ER_mod)
pnld_e_10_ER_mod$coefficients[2]
pnld_e_10_ER_slope <- unname(pnld_e_10_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_10_ER <- pnld_e_10_ER_slope * 1440
pnld_e_10_ER

######################## GPP
pnld_e_10_GPP <- pnld_e_10_NEP - pnld_e_10_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_10_plot <- dat_full %>%
  filter(chamber == "10") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_10_NEP_mod$coefficients[1], slope = pnld_e_10_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_10_ER_mod$coefficients[1], slope = pnld_e_10_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_10_plot
```

#### Chamber 12
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_12 <- dat_full %>%
  filter(chamber == "12" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_12_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_12)
summary(pnld_e_12_NEP_mod)
pnld_e_12_NEP_mod$coefficients[2]
pnld_e_12_NEP_slope <- unname(pnld_e_12_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_12_NEP <- pnld_e_12_NEP_slope * 1440
pnld_e_12_NEP

######################## ER rate calculations

pnld_e_12 <- dat_full %>%
  filter(chamber == "12" & delta_T_min >= 1206)

pnld_e_12_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_12)
summary(pnld_e_12_ER_mod)
pnld_e_12_ER_mod$coefficients[2]
pnld_e_12_ER_slope <- unname(pnld_e_12_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_12_ER <- pnld_e_12_ER_slope * 1440
pnld_e_12_ER

######################## GPP
pnld_e_12_GPP <- pnld_e_12_NEP - pnld_e_12_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_12_plot <- dat_full %>%
  filter(chamber == "12") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_12_NEP_mod$coefficients[1], slope = pnld_e_12_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_12_ER_mod$coefficients[1], slope = pnld_e_12_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_12_plot
```
##### R2 below 0.95

#### Chamber 13
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_13 <- dat_full %>%
  filter(chamber == "13" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_13_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_13)
summary(pnld_e_13_NEP_mod)
pnld_e_13_NEP_mod$coefficients[2]
pnld_e_13_NEP_slope <- unname(pnld_e_13_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_13_NEP <- pnld_e_13_NEP_slope * 1440
pnld_e_13_NEP

######################## ER rate calculations

pnld_e_13 <- dat_full %>%
  filter(chamber == "13" & delta_T_min >= 1206)

pnld_e_13_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_13)
summary(pnld_e_13_ER_mod)
pnld_e_13_ER_mod$coefficients[2]
pnld_e_13_ER_slope <- unname(pnld_e_13_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_13_ER <- pnld_e_13_ER_slope * 1440
pnld_e_13_ER

######################## GPP
pnld_e_13_GPP <- pnld_e_13_NEP - pnld_e_13_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_13_plot <- dat_full %>%
  filter(chamber == "13") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_13_NEP_mod$coefficients[1], slope = pnld_e_13_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_13_ER_mod$coefficients[1], slope = pnld_e_13_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_13_plot
```
##### R2 below 0.95

### Tank 1; 12.5C
#### Chamber 3
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_3 <- dat_full %>%
  filter(chamber == "3" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_3_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_3)
summary(pnld_e_3_NEP_mod)
pnld_e_3_NEP_mod$coefficients[2]
pnld_e_3_NEP_slope <- unname(pnld_e_3_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_3_NEP <- pnld_e_3_NEP_slope * 1440
pnld_e_3_NEP

######################## ER rate calculations

pnld_e_3 <- dat_full %>%
  filter(chamber == "3" & delta_T_min >= 1206)

pnld_e_3_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_3)
summary(pnld_e_3_ER_mod)
pnld_e_3_ER_mod$coefficients[2]
pnld_e_3_ER_slope <- unname(pnld_e_3_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_3_ER <- pnld_e_3_ER_slope * 1440
pnld_e_3_ER

######################## GPP
pnld_e_3_GPP <- pnld_e_3_NEP - pnld_e_3_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_3_plot <- dat_full %>%
  filter(chamber == "3") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_3_NEP_mod$coefficients[1], slope = pnld_e_3_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_3_ER_mod$coefficients[1], slope = pnld_e_3_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_3_plot
```
##### R2 below 0.95

#### Chamber 7
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_7 <- dat_full %>%
  filter(chamber == "7" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_7_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_7)
summary(pnld_e_7_NEP_mod)
pnld_e_7_NEP_mod$coefficients[2]
pnld_e_7_NEP_slope <- unname(pnld_e_7_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_7_NEP <- pnld_e_7_NEP_slope * 1440
pnld_e_7_NEP

######################## ER rate calculations

pnld_e_7 <- dat_full %>%
  filter(chamber == "7" & delta_T_min >= 1206)

pnld_e_7_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_7)
summary(pnld_e_7_ER_mod)
pnld_e_7_ER_mod$coefficients[2]
pnld_e_7_ER_slope <- unname(pnld_e_7_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_7_ER <- pnld_e_7_ER_slope * 1440
pnld_e_7_ER

######################## GPP
pnld_e_7_GPP <- pnld_e_7_NEP - pnld_e_7_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_7_plot <- dat_full %>%
  filter(chamber == "7") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_7_NEP_mod$coefficients[1], slope = pnld_e_7_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_7_ER_mod$coefficients[1], slope = pnld_e_7_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_7_plot
```
##### R2 below 0.95

#### Chamber 15
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_15 <- dat_full %>%
  filter(chamber == "15" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_15_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_15)
summary(pnld_e_15_NEP_mod)
pnld_e_15_NEP_mod$coefficients[2]
pnld_e_15_NEP_slope <- unname(pnld_e_15_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_15_NEP <- pnld_e_15_NEP_slope * 1440
pnld_e_15_NEP

######################## ER rate calculations

pnld_e_15 <- dat_full %>%
  filter(chamber == "15" & delta_T_min >= 1206)

pnld_e_15_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_15)
summary(pnld_e_15_ER_mod)
pnld_e_15_ER_mod$coefficients[2]
pnld_e_15_ER_slope <- unname(pnld_e_15_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_15_ER <- pnld_e_15_ER_slope * 1440
pnld_e_15_ER

######################## GPP
pnld_e_15_GPP <- pnld_e_15_NEP - pnld_e_15_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_15_plot <- dat_full %>%
  filter(chamber == "15") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_15_NEP_mod$coefficients[1], slope = pnld_e_15_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_15_ER_mod$coefficients[1], slope = pnld_e_15_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_15_plot
```
##### R2 below 0.95

#### Chamber 16
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_16 <- dat_full %>%
  filter(chamber == "16" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_16_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_16)
summary(pnld_e_16_NEP_mod)
pnld_e_16_NEP_mod$coefficients[2]
pnld_e_16_NEP_slope <- unname(pnld_e_16_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_16_NEP <- pnld_e_16_NEP_slope * 1440
pnld_e_16_NEP

######################## ER rate calculations

pnld_e_16 <- dat_full %>%
  filter(chamber == "16" & delta_T_min >= 1206)

pnld_e_16_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_16)
summary(pnld_e_16_ER_mod)
pnld_e_16_ER_mod$coefficients[2]
pnld_e_16_ER_slope <- unname(pnld_e_16_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_16_ER <- pnld_e_16_ER_slope * 1440
pnld_e_16_ER

######################## GPP
pnld_e_16_GPP <- pnld_e_16_NEP - pnld_e_16_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_16_plot <- dat_full %>%
  filter(chamber == "16") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_16_NEP_mod$coefficients[1], slope = pnld_e_16_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_16_ER_mod$coefficients[1], slope = pnld_e_16_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_16_plot
```
##### R2 below 0.95

### Tank 3; 15.5C
#### Chamber 1
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_1 <- dat_full %>%
  filter(chamber == "1" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_1_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_1)
summary(pnld_e_1_NEP_mod)
pnld_e_1_NEP_mod$coefficients[2]
pnld_e_1_NEP_slope <- unname(pnld_e_1_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_1_NEP <- pnld_e_1_NEP_slope * 1440
pnld_e_1_NEP

######################## ER rate calculations

pnld_e_1 <- dat_full %>%
  filter(chamber == "1" & delta_T_min >= 1225)

pnld_e_1_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_1)
summary(pnld_e_1_ER_mod)
pnld_e_1_ER_mod$coefficients[2]
pnld_e_1_ER_slope <- unname(pnld_e_1_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_1_ER <- pnld_e_1_ER_slope * 1440
pnld_e_1_ER

######################## GPP
pnld_e_1_GPP <- pnld_e_1_NEP - pnld_e_1_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_1_plot <- dat_full %>%
  filter(chamber == "1") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_1_NEP_mod$coefficients[1], slope = pnld_e_1_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_1_ER_mod$coefficients[1], slope = pnld_e_1_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_1_plot
```
##### R2 below 0.95

#### Chamber 2
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_2 <- dat_full %>%
  filter(chamber == "2" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_2_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_2)
summary(pnld_e_2_NEP_mod)
pnld_e_2_NEP_mod$coefficients[2]
pnld_e_2_NEP_slope <- unname(pnld_e_2_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_2_NEP <- pnld_e_2_NEP_slope * 1440
pnld_e_2_NEP

######################## ER rate calculations

pnld_e_2 <- dat_full %>%
  filter(chamber == "2" & delta_T_min >= 1225)

pnld_e_2_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_2)
summary(pnld_e_2_ER_mod)
pnld_e_2_ER_mod$coefficients[2]
pnld_e_2_ER_slope <- unname(pnld_e_2_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_2_ER <- pnld_e_2_ER_slope * 1440
pnld_e_2_ER

######################## GPP
pnld_e_2_GPP <- pnld_e_2_NEP - pnld_e_2_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_2_plot <- dat_full %>%
  filter(chamber == "2") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_2_NEP_mod$coefficients[1], slope = pnld_e_2_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_2_ER_mod$coefficients[1], slope = pnld_e_2_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_2_plot
```

#### Chamber 4
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_4 <- dat_full %>%
  filter(chamber == "4" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_4_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_4)
summary(pnld_e_4_NEP_mod)
pnld_e_4_NEP_mod$coefficients[2]
pnld_e_4_NEP_slope <- unname(pnld_e_4_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_4_NEP <- pnld_e_4_NEP_slope * 1440
pnld_e_4_NEP

######################## ER rate calculations

pnld_e_4 <- dat_full %>%
  filter(chamber == "4" & delta_T_min >= 1225)

pnld_e_4_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_4)
summary(pnld_e_4_ER_mod)
pnld_e_4_ER_mod$coefficients[2]
pnld_e_4_ER_slope <- unname(pnld_e_4_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_4_ER <- pnld_e_4_ER_slope * 1440
pnld_e_4_ER

######################## GPP
pnld_e_4_GPP <- pnld_e_4_NEP - pnld_e_4_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_4_plot <- dat_full %>%
  filter(chamber == "4") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_4_NEP_mod$coefficients[1], slope = pnld_e_4_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_4_ER_mod$coefficients[1], slope = pnld_e_4_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_4_plot
```

#### Chamber 14
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_14 <- dat_full %>%
  filter(chamber == "14" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_14_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_14)
summary(pnld_e_14_NEP_mod)
pnld_e_14_NEP_mod$coefficients[2]
pnld_e_14_NEP_slope <- unname(pnld_e_14_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_14_NEP <- pnld_e_14_NEP_slope * 1440
pnld_e_14_NEP

######################## ER rate calculations

pnld_e_14 <- dat_full %>%
  filter(chamber == "14" & delta_T_min >= 1225)

pnld_e_14_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_14)
summary(pnld_e_14_ER_mod)
pnld_e_14_ER_mod$coefficients[2]
pnld_e_14_ER_slope <- unname(pnld_e_14_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_14_ER <- pnld_e_14_ER_slope * 1440
pnld_e_14_ER

######################## GPP
pnld_e_14_GPP <- pnld_e_14_NEP - pnld_e_14_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_14_plot <- dat_full %>%
  filter(chamber == "14") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_14_NEP_mod$coefficients[1], slope = pnld_e_14_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_14_ER_mod$coefficients[1], slope = pnld_e_14_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_14_plot
```

### Tank 4; 18.5C
#### Chamber 6
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_6 <- dat_full %>%
  filter(chamber == "6" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_6_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_6)
summary(pnld_e_6_NEP_mod)
pnld_e_6_NEP_mod$coefficients[2]
pnld_e_6_NEP_slope <- unname(pnld_e_6_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_6_NEP <- pnld_e_6_NEP_slope * 1440
pnld_e_6_NEP

######################## ER rate calculations

pnld_e_6 <- dat_full %>%
  filter(chamber == "6" & delta_T_min >= 1206)

pnld_e_6_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_6)
summary(pnld_e_6_ER_mod)
pnld_e_6_ER_mod$coefficients[2]
pnld_e_6_ER_slope <- unname(pnld_e_6_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_6_ER <- pnld_e_6_ER_slope * 1440
pnld_e_6_ER

######################## GPP
pnld_e_6_GPP <- pnld_e_6_NEP - pnld_e_6_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_6_plot <- dat_full %>%
  filter(chamber == "6") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_6_NEP_mod$coefficients[1], slope = pnld_e_6_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_6_ER_mod$coefficients[1], slope = pnld_e_6_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_6_plot
```

#### Chamber 8
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_8 <- dat_full %>%
  filter(chamber == "8" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_8_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_8)
summary(pnld_e_8_NEP_mod)
pnld_e_8_NEP_mod$coefficients[2]
pnld_e_8_NEP_slope <- unname(pnld_e_8_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_8_NEP <- pnld_e_8_NEP_slope * 1440
pnld_e_8_NEP

######################## ER rate calculations

pnld_e_8 <- dat_full %>%
  filter(chamber == "8" & delta_T_min >= 1206)

pnld_e_8_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_8)
summary(pnld_e_8_ER_mod)
pnld_e_8_ER_mod$coefficients[2]
pnld_e_8_ER_slope <- unname(pnld_e_8_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_8_ER <- pnld_e_8_ER_slope * 1440
pnld_e_8_ER

######################## GPP
pnld_e_8_GPP <- pnld_e_8_NEP - pnld_e_8_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_8_plot <- dat_full %>%
  filter(chamber == "8") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_8_NEP_mod$coefficients[1], slope = pnld_e_8_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_8_ER_mod$coefficients[1], slope = pnld_e_8_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_8_plot
```

#### Chamber 9
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_9 <- dat_full %>%
  filter(chamber == "9" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_9_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_9)
summary(pnld_e_9_NEP_mod)
pnld_e_9_NEP_mod$coefficients[2]
pnld_e_9_NEP_slope <- unname(pnld_e_9_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_9_NEP <- pnld_e_9_NEP_slope * 1440
pnld_e_9_NEP

######################## ER rate calculations

pnld_e_9 <- dat_full %>%
  filter(chamber == "9" & delta_T_min >= 1206)

pnld_e_9_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_9)
summary(pnld_e_9_ER_mod)
pnld_e_9_ER_mod$coefficients[2]
pnld_e_9_ER_slope <- unname(pnld_e_9_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_9_ER <- pnld_e_9_ER_slope * 1440
pnld_e_9_ER

######################## GPP
pnld_e_9_GPP <- pnld_e_9_NEP - pnld_e_9_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_9_plot <- dat_full %>%
  filter(chamber == "9") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_9_NEP_mod$coefficients[1], slope = pnld_e_9_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_9_ER_mod$coefficients[1], slope = pnld_e_9_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_9_plot
```
##### R2 below 0.95

#### Chamber 11
```{r}
######################## NEP rate calculations

## the linear portion of the NEP curve is roughly from minute 1117 to minute 1206 ##

pnld_e_11 <- dat_full %>%
  filter(chamber == "11" & delta_T_min >= 1117 & delta_T_min <= 1206)

pnld_e_11_NEP_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_11)
summary(pnld_e_11_NEP_mod)
pnld_e_11_NEP_mod$coefficients[2]
pnld_e_11_NEP_slope <- unname(pnld_e_11_NEP_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_11_NEP <- pnld_e_11_NEP_slope * 1440
pnld_e_11_NEP

######################## ER rate calculations

pnld_e_11 <- dat_full %>%
  filter(chamber == "11" & delta_T_min >= 1206)

pnld_e_11_ER_mod <- lm(oxygen_mass_mg~delta_T_min, pnld_e_11)
summary(pnld_e_11_ER_mod)
pnld_e_11_ER_mod$coefficients[2]
pnld_e_11_ER_slope <- unname(pnld_e_11_ER_mod$coefficients[2])


## Now converting this rate from mg O2 per minute to per day by multiplying this rate by 1440 minutes per day. ##
pnld_e_11_ER <- pnld_e_11_ER_slope * 1440
pnld_e_11_ER

######################## GPP
pnld_e_11_GPP <- pnld_e_11_NEP - pnld_e_11_ER

# plot of DO in this chamber to check that regressions match with data
pnld_e_11_plot <- dat_full %>%
  filter(chamber == "11") %>%
  ggplot(aes(x = delta_T_min, y = oxygen_mass_mg)) +
  geom_point(shape = 1, size = 2) +
  geom_abline(aes(intercept = pnld_e_11_NEP_mod$coefficients[1], slope = pnld_e_11_NEP_mod$coefficients[2])) +
  geom_abline(aes(intercept = pnld_e_11_ER_mod$coefficients[1], slope = pnld_e_11_ER_mod$coefficients[2])) +
  theme_classic()
pnld_e_11_plot
```


## Creating dataframe with metab rates
```{r}
# creating dataframe of metab rates to make barplot
pnld_e_rates <- data.frame(
  "NEP_mg_d" = c(pnld_e_1_NEP, pnld_e_2_NEP, pnld_e_3_NEP, pnld_e_4_NEP, pnld_e_5_NEP, pnld_e_6_NEP, pnld_e_7_NEP, pnld_e_8_NEP, pnld_e_9_NEP, pnld_e_10_NEP, pnld_e_11_NEP, pnld_e_12_NEP, pnld_e_13_NEP, pnld_e_14_NEP, pnld_e_15_NEP, pnld_e_16_NEP),
  "ER_mg_d" = c(pnld_e_1_ER, pnld_e_2_ER, pnld_e_3_ER, pnld_e_4_ER, pnld_e_5_ER, pnld_e_6_ER, pnld_e_7_ER, pnld_e_8_ER, pnld_e_9_ER, pnld_e_10_ER, pnld_e_11_ER, pnld_e_12_ER, pnld_e_13_ER, pnld_e_14_ER, pnld_e_15_ER,  pnld_e_16_ER),
  "chamber" = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16"))
head(pnld_e_rates)
#str(pnld_e_rates)

# formatting data types of columns
pnld_e_rates$NEP_mg_d <- as.numeric(pnld_e_rates$NEP_mg_d)
pnld_e_rates$ER_mg_d <- as.numeric(pnld_e_rates$ER_mg_d)

# calculating GPP from NEP and ER columns
pnld_e_rates <- pnld_e_rates %>%
  mutate(GPP_mg_d = NEP_mg_d - ER_mg_d)
head(pnld_e_rates)
```

## Joining metab rates dataframe with chamber dataframe
```{r}
# changing chamber column in datasheet to character to be compatable to joining to pnld_a_rates
dat_chambers$chamber <- as.factor(dat_chambers$chamber)

# joining chamber and metabolic rate data
dat_final <- full_join(dat_chambers, pnld_e_rates, by = "chamber")
#View(dat_final)

# adding temperature treatment (discrete factor) and temp_C (continuous numeric) columns based off of tank column already in data
dat_final <- dat_final %>%
  mutate(
    treatment = case_when(
      Tank_e == "1" ~ "12.5C",
      Tank_e == "2" ~ "9.5C",
      Tank_e == "3" ~ "15.5C",
      Tank_e == "4" ~ "18.5C"
    )) %>%
  mutate(
    temp_C = case_when(
           Tank_e == "1" ~ 12.5,
           Tank_e == "2" ~ 9.5,
           Tank_e == "3" ~ 15.5,
           Tank_e == "4" ~ 18.5
           ))


#View(dat_final)
```

## Normalizing metabolic rates
```{r}
# normalizing metabolic rates by AFDW and by rock surface area
dat_final <- dat_final %>%
  mutate(GPP_mgO2_d_gAFDW = GPP_mg_d/total_AFDW_final, NEP_mgO2_d_gAFDW = NEP_mg_d/total_AFDW_final, ER_mgO2_d_gAFDW = ER_mg_d/total_AFDW_final) %>%
  mutate(GPP_mgO2_d_m2 = GPP_mg_d/rock_SA_m2, NEP_mgO2_d_m2 = NEP_mg_d/rock_SA_m2, ER_mgO2_d_m2 = ER_mg_d/rock_SA_m2)

#View(dat_final)
```

# Exporting metabolic rate data for further analysis
```{r}
# making identifier column for joining to ambient incubation data for nutrient effect analysis
dat_final <- dat_final %>%
  mutate(nutrient_trt = "enriched")
# exporting data for final analysis
write_csv(dat_final, "data/211128_pineland_a_e/211128_pnld_enriched_rates.csv")
```