---
title: "Peri_biomass_plot_adapted_from_KA2021"
output: html_document
date: "2022-12-22"
---
Creating plot adapted from figure 4 in atkins et al 2021 (DOI: 10.1007/s10750-021-04533-w) of periphyton biomass in Lake Tahoe.

Input data files include median periphyton biomass values (AFDW and Chla) between 1982 and 2019 that were extracted from the paper using webplot digitizer on 12/22/22 (months coded as numbers (1-12 starting with sept)).

Output files are plots of these values exported as jpg files.

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)
library(arm)
library(nlme)
library(lattice)
library(emmeans)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(patchwork)
library(gt)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_chunk$set(root.dir = '..')
```

# Loading data frame
```{r, include = FALSE}
# importing AFDW data
### colnames are not in the original file, so they're added below
AFDW <- read_csv("data/main_analysis/Karen_Atkins_AFDW_data.csv", col_names = FALSE)
colnames(AFDW) <- c("Month_num", "AFDW_g_m2")
#View(AFDW)

# importing Chla data
### colnames are not in the original file, so they're added below
Chla <- read_csv("data/main_analysis/Karen_Atkins_Chla_data.csv", col_names = FALSE)
colnames(Chla) <- c("Month_num", "Chla_mg_m2")
#View(Chla)

# creating vector of months (with correct levels) to combine with the above data frames (goes by water year Oct-Oct)
Month <- c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep")
Month = factor(Month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# appending month column to dataframe
AFDW <- data.frame(AFDW, Month)

Chla <- data_frame(Chla, Month)

# adjusting the order of levels so that they plot in the calendar-year order
Month = factor(Month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

```


# AFDW plot
```{r}
AFDW_plot <- AFDW %>%
  ggplot(aes(x = Month, y = AFDW_g_m2, group = 1)) +
  geom_point(size = 3) +
  scale_y_continuous(limits = c(0,30))+
  geom_line(linetype = 2, linewidth = 1) +
  labs(y = expression(AFDW~(g/m^{2})), x = "") +
  theme_classic()
AFDW_plot

Chla_plot <- Chla %>%
  ggplot(aes(x = Month, y = Chla_mg_m2, group = 1)) +
  geom_point(size = 3) +
  scale_y_continuous(limits = c(0,33))+
  geom_line(linetype = 2, linewidth = 1) +
  labs(y = expression(Chl-a~(mg/m^{2}))) +
  theme_classic()
Chla_plot
```

# Combining plots
```{r}
KA_biomass_trends <- AFDW_plot/Chla_plot

ggsave("plots/main_analysis/KA_biomass_trends.jpg", plot = KA_biomass_trends, height = 6, width = 6, dpi = 300)

ggsave("plots/final_paper/Figure1_inset.jpg", plot = KA_biomass_trends, height = 6, width = 6, dpi = 300)
```


