---
title: "Tahoe_peri_data_prep_explor_stats"
output: html_document
date: "2023-06-19"
---
Input data files: cleaned dataset generated in "Tahoe_Peri_data_joining.Rmd" script.

Output data files: Descriptive plots of background conditions measured in the field and exploratory plots of relationships between periphyton metabolism and other variables.

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)
library(arm)
library(nlme)
library(lattice)
library(emmeans)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(patchwork)
library(gt)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_chunk$set(root.dir = '..')
```

# importing data
```{r}
# dataframe with metabolic rate data, but preliminary experiment data is excluded
dat_clean <- read_csv("data/main_analysis/cleaned_rate_data.csv")

# dataframe with metabolic rate data from all experiments
dat <- read_csv("data/main_analysis/main_rate_data.csv")

# data
main_dat <- read_csv("data/main_analysis/main_data_lahontan.csv")
```

# Plotting main data
## Comparing AFDW: original vs. corrected
```{r}
AFDW_plot <- dat_clean %>%
  ggplot(aes(x = total_AFDW_g, y = total_AFDW_g_2)) +
  geom_point() +
  geom_abline(aes(intercept = 0, slope = 1)) +
  geom_rug(sides = "bl") +
  labs(x = "Old AFDW", y = "Corrected AFDW", title = "Comparison of AFDW Data") +
  lims( x = c(0, 1.5), y = c(0, 1.5)) +
  theme_classic()
AFDW_plot

ggsave("plots/main_analysis/AFDW_comparison.png", plot = AFDW_plot, width = 3, height = 3)
```

## New AFDW-adjusted Metab. rate plot
### Figure S4
```{r}
# remelting dataframe to long format so that metabolic rate columns are combined into one column and values are in a separate column so all can be plotted at once
dat_long <- dat_clean %>%
  rename(GPP = GPP_mgO2_d_gAFDW_2, ER = ER_mgO2_d_gAFDW_2, NEP = NEP_mgO2_d_gAFDW_2) %>% # renaming columns so that facet labels are more tidy
  pivot_longer(cols = c(NEP, GPP, ER),
               names_to = "rate_name_AFDW",
               values_to = "rates_AFDW",
               values_drop_na = TRUE)
#View(dat_long)

# making experiment ID column a factor instead of numeric since it is treated as discrete rather than continuous in the plot
dat_long$exp_ID <- as.factor(dat_long$exp_ID)


AFDW_2_rates <- dat_long %>%
  filter(month != "april") %>% 
  ggplot(aes(x = temp_C, y = rates_AFDW/24, group = interaction(exp_ID, nutrient_trt))) +
  facet_grid(cols = vars(rate_name_AFDW),
             rows = vars(nutrient_trt)) +
  #scale_y_log10() +
 geom_point(aes(shape = month,
                #color = nutrient_trt
                ), size = 1.5) +
  #geom_smooth(method = "lm", se = FALSE) +
  #scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Month") +# renaming label of point shape legend
  #scale_shape_manual(values = c(15:17, 3:5, 8)) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~hr^{-1}~gAFDW^{-1})))

AFDW_2_rates

# high quality dpi plot
ggsave("plots/main_analysis/correct_AFDW_norm_rates.png", plot = AFDW_2_rates, width = 5, height = 5, dpi = 300)

ggsave("plots/final_paper/FigureS4.jpg", plot = AFDW_2_rates, width = 5, height = 5, dpi = 300)

```

## Plotting Areal Metab. Rates
```{r}
# melting dataframe to long format so that metabolic rate columns are combined into one column and values are in a separate column so all can be plotted at once
dat_long_m2 <- dat_clean %>%
  rename(GPP = GPP_mgO2_d_m2, ER = ER_mgO2_d_m2, NEP = NEP_mgO2_d_m2) %>% # renaming columns so that facet labels are more tidy
  pivot_longer(cols = c(NEP, GPP, ER),
               names_to = "rate_name_m2",
               values_to = "rates_m2",
               values_drop_na = TRUE)
#View(dat_long_m2)

# making experiment ID column a factor instead of numeric since it is treated as discrete rather than continuous in the plot
dat_long_m2$exp_ID <- as.factor(dat_long_m2$exp_ID)

areal_rates_log_log_AGU <- dat_long_m2 %>%
  ggplot(aes(x = temp_C, y = rates_m2, group = interaction(exp_ID, nutrients))) +
  facet_grid(cols = vars(rate_name_m2)) +
  scale_y_log10() +
 geom_point(aes(shape = exp_ID, color = nutrient_trt), size = 1.5) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Experiment ID") +# renaming label of point shape legend
  scale_shape_manual(values = c(0:6)) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~m^{-2})))

areal_rates_log_log_AGU

ggsave("plots/main_analysis/areal_rates_unlogged_AGU.png", plot = areal_rates_log_log_AGU, width = 7, height = 3, dpi = 300)

# same plot, but removing april experiments since they were excluded from models (due to different experimental design)

areal_rates_log_wo_april <- dat_long_m2 %>%
  filter(month != "april") %>%
  ggplot(aes(x = temp_C, y = rates_m2, group = interaction(exp_ID, nutrients))) +
  facet_grid(cols = vars(rate_name_m2)) +
  scale_y_log10() +
 geom_point(aes(shape = exp_ID, color = nutrient_trt), size = 1.5) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_manual(name = "Experiment ID", values = c(15:17, 3:5, 8)) + # renaming label of point shape legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~m^{-2})))

areal_rates_log_wo_april

ggsave("plots/main_analysis/areal_rates_unlogged_wo_april.png", plot = areal_rates_log_wo_april, width = 8, height = 3, dpi = 300)



# same plot, but ungrouping by experiment ID and adding back in april 2021 data

areal_rates <- dat_long_m2 %>%
  ggplot(aes(x = temp_C, y = rates_m2, group = nutrients)) +
  facet_grid(cols = vars(rate_name_m2)) +
  scale_y_log10() +
 geom_point(aes(color = nutrient_trt), size = 1.5) +
  geom_smooth(aes(color = nutrient_trt), method = "lm", se = FALSE) +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~m^{-2})))

areal_rates

ggsave("plots/main_analysis/SA_model_output.png", plot = areal_rates, width = 7, height = 3, dpi = 300)
```

## Chl-a-normalized GPP plot
```{r}
# normalizing GPP rates by mass of chla on rocks
chla_rates <- dat_clean %>%
  mutate(GPP_mgO2_d_ugChla = GPP_mg_d/(total_chla_ug/1e6)) #converting chla data from ug to g

# making experiment ID column a factor instead of numeric since it is treated as discrete rather than continuous in the plot
chla_rates$exp_ID <- as.factor(chla_rates$exp_ID)

# now plotting GPP rates vs. temp, and removing april 2021 data

GPP_chla <- chla_rates %>%
  filter(month != "april") %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_ugChla, group = interaction(exp_ID, nutrients))) +
  scale_y_log10() +
 geom_point(aes(shape = exp_ID, color = nutrient_trt), size = 1.5) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_manual(name = "Experiment ID", values = c(15:17, 3:5, 8)) + # renaming label of point shape legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gChla^{-1})))

GPP_chla

ggsave("plots/main_analysis/GPP_chla.png", plot = GPP_chla, height = 4, width = 8)


```

## Autotrophic Index Seasonal Patterns
```{r}
# Using the autotrophic index (ratio of AFDW:chl-a) (Biggs & Close 1989)
         
dat %>%
  group_by(date) %>%
  summarise(AI_avg = mean(AI), AI_sd = sd(AI)) %>%
  arrange(AI_avg)

```

# Temperature Randomization Checks
Checking that randomization of temperature treatments between ambient and enriched nutrient trials did not have any drastic effects on metabolic rates
```{r}
temp_diff <- dat_clean %>% 
  group_by(month, exp_ID, rock_ID) %>% 
  arrange(month, exp_ID, rock_ID, nutrient_trt) %>% 
  summarise(temp_difference = temp_C - lag(temp_C),
            GPP_diff = GPP_mg_d - lag(GPP_mg_d),
            ER_diff = ER_mgO2_d_m2 - lag(ER_mgO2_d_m2),
            NEP_diff = NEP_mg_d - lag(NEP_mg_d)) %>% 
  drop_na() %>% 
  pivot_longer(GPP_diff:NEP_diff, names_to = "parameter", values_to = "value")

temp_diff %>% 
  mutate(abs_difference = abs(value)) %>% 
  ggplot(aes(x = factor(temp_difference), y = value,
             #col = month
             ))+
  facet_wrap(vars(month, parameter),
             nrow = 5,
             scales = "free_y")+
  geom_boxplot()+
  geom_hline(yintercept = 0, linetype = 3)+
  theme_bw()+
  labs(x = "Temperature Difference (C)",
       y = "Metabolic Rate Absolute Difference")
```


# Exploratory Data Analysis
```{r}
# removing data from march experiments since it was at a different site and had excessive variation in AFDW-normalized GPP rates due to a discrepancy in filtering methods
dat_pnld <- dat %>%
  filter(month != "march")

# histogram of temperature data to check normality
dat_pnld %>%
  ggplot(aes(temp_C)) +
  geom_histogram()
# looks approximately like a normal distribution

# now checking normality in AFDW-normalized metabolic rates
dat_pnld %>%
  ggplot(aes(GPP_mgO2_d_gAFDW)) +
  geom_histogram()

dat_pnld %>%
  ggplot(aes(NEP_mgO2_d_gAFDW)) +
  geom_histogram()

dat_pnld %>%
  ggplot(aes(ER_mgO2_d_gAFDW)) +
  geom_histogram()
# some definite outliers present, also some potential for a right-skewed distn in the ER and GPP rates.

# cleveland plot
# dotchart(dat_pnld$)

# selecting only the numeric columns for calculating a correlation matrix
dat_cor <- dat_pnld %>%
  dplyr::select(temp_C, GPP_mgO2_d_gAFDW, NEP_mgO2_d_gAFDW, ER_mgO2_d_gAFDW, GPP_mgO2_d_m2, NEP_mgO2_d_m2, ER_mgO2_d_m2, NO3_N_ug_L, NH4_N_ug_L, SRP_ug_L)

# standardizing data with mean = 0, sd = 1 for correlation matrix
dat_cor <- scale(dat_cor)
#View(dat_cor)

# plotting correlation matrix of the data
cor(dat_cor, use = "complete.obs")

```

# Time-series plots
### chla
```{r}
### chlorophyll-a trends across time

# making bounds for x axis in plot below
min <- as.Date("2021-04-01")
max <- as.Date("2022-04-01")


chla_trends <- dat %>%
  filter(exp_ID != 1 & exp_ID != 2) %>% # filtering out march 1st and 15th data since they were performed at a different site
  mutate(chla_g_m2 = total_chla_ug/rock_SA_m2/1e6) %>% # calculating areal chlorophyll-a concentrations (per m2)
  ggplot(aes(x = date, y = chla_g_m2, group = date)) +
  geom_boxplot() +
  labs(x = "", y = expression(Periphyton~Chl-a~(g~m^{-2}))) +
  scale_x_date(limits = c(min, max)) +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

chla_trends

# simplified plot w/o expressions for ggplotly

chla_plotly <- dat %>%
  filter(exp_ID != 1) %>% # filtering out march 1st data due to different chla sampling protocols used for this experiment
  mutate(chla_g_m2 = total_chla_ug/rock_SA_m2/1e6) %>% # calculating areal chlorophyll-a concentrations (per m2)
  ggplot(aes(x = date, y = chla_g_m2, group = date)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p <- plotly::ggplotly(chla_plotly)
p
```

### AI
```{r}
# plotting changes in AI over the year

AI_trends <- dat %>%
  filter(exp_ID != 1 & exp_ID != 2) %>% # filtering out march 1st and 15th data since they were performed at a different site
  filter(AI < 5000) %>% # filtering out 2 outliers from october and march
  ggplot(aes(x = date, y = AI, group = date)) +
  geom_boxplot() +
  labs(x = "", y = "Autotrophic Index (gAFDW:gChla)") +
  scale_x_date(limits = c(min, max)) +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

AI_trends

ggsave("plots/main_analysis/AI_timeseries.png", plot = AI_trends, height = 4, width = 8, dpi = 300)
```

### AFDW
```{r}
# plotting changes in areal AFDW over the year

AFDW_trends <- dat %>%
  filter(exp_ID != 1 & exp_ID != 2) %>% # filtering out march 1st and 15th data since they were performed at a different site
  mutate(AFDW_g_m2 = total_AFDW_g_2/rock_SA_m2) %>%
  filter(AFDW_g_m2 <= 200) %>% # filtering out outlier from exp_ID 2
  ggplot(aes(x = date, y = AFDW_g_m2, group = date)) +
  geom_boxplot() +
  labs(x = "", y = expression(Periphyton~AFDW~(g~m^{-2}))) +
  scale_x_date(limits = c(min, max)) +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

AFDW_trends

ggsave("plots/main_analysis/AFDW_timeseries.png", plot = AFDW_trends, height = 4, width = 8, dpi = 300)


# simplified plot for plotly
AFDW_plotly <- dat %>%
  filter(exp_ID != 1) %>% # filtering out march 1st data due to different chla sampling protocols used for this experiment
  mutate(AFDW_g_m2 = total_AFDW_g_2/rock_SA_m2) %>%
  filter(AFDW_g_m2 <= 200) %>% # filtering out outlier from exp_ID 2
  ggplot(aes(x = date, y = AFDW_g_m2, group = date)) +
  geom_boxplot() +
  theme_classic()

p <- plotly::ggplotly(AFDW_plotly)
p

```

### Temp
#### Figure 2
```{r}
# making bounds for x axis in plot below
min <- as.Date("2021-04-01")
max <- as.Date("2022-04-01")

# plotting in-situ temps for each experiment
insitu_temp <- dat %>%
 filter(exp_ID != 1 & exp_ID != 2) %>% # filtering out march 1st and 15th data since they were performed at a different site
  filter(exp_ID == 2 & temp_C == 7 | exp_ID == 3 & temp_C == 7 | exp_ID == 4 & temp_C == 7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only data from in-situ temperatures in the field
  ggplot(aes( x = date, y = temp_C)) +
  geom_point(size = 3) +
  geom_line() +
  labs(y = expression('Temperature (\u00B0C)'), x = "Date") +
  scale_x_date(date_labels = "%b %Y", limits = c(min, max)) +
  theme_classic()
insitu_temp

# patching together AI, AFDW, chl-a, and in situ temp timeseries plots so that they share the same x-axis
library(patchwork)

AI_AFDW <- (chla_trends/AFDW_trends/AI_trends/insitu_temp) + plot_annotation(tag_levels = 'a')

ggsave("plots/main_analysis/AI_AFDW_trends.jpg", plot = AI_AFDW, height = 11, width = 11, dpi = 300)

ggsave("plots/final_paper/Figure2.jpg", plot = AI_AFDW, height = 10, width = 8, dpi = 300)
```


### P deficiency index (Lean & Pick 1981)
```{r}
# calculating phosphorus deficiency index of periphyton as in Lean & Pick 1981 and using this to assess nutrient limitation in periphyton communities

# plotting NEP rates vs. SRP uptake rates to get an idea of the dispersion pattern
main_dat %>%
  filter(!is.na(SRP_rate) & SRP_rate <= 200) %>% # filtering out rows with no SRP uptake rate and a few outlier rows
  ggplot(aes(x = SRP_rate, y = NEP_mgO2_d_gAFDW_2, color = month)) +
  geom_point() +
  labs(x = expression(PO[4]~Uptake~(ugP~hr^{-1}~gAFDW^{-1})), y = expression(NEP~(mgO[2]~day^{-1}~gAFDW^{-1})))

# making same plot, but averaging replicates by temperature and month

Pdef_plot <- main_dat %>%
  filter(!is.na(SRP_rate) & SRP_rate <= 200 & exp_ID >= 4) %>% # filtering out rows with no SRP uptake rate, a few outlier rows, and rows from march 2021 and April 7 2021 experiments
  group_by(temp_C, date) %>% # grouping dataframe by temperature value and month
  summarise(month, avgSRP = mean(SRP_rate), lowSRP = mean(SRP_rate) - sd(SRP_rate), upperSRP = mean(SRP_rate) + sd(SRP_rate), avgNEP = mean(NEP_mgO2_d_gAFDW_2), lowNEP = mean(NEP_mgO2_d_gAFDW_2) - sd(NEP_mgO2_d_gAFDW_2), upperNEP = mean(NEP_mgO2_d_gAFDW_2) + sd(NEP_mgO2_d_gAFDW_2)) %>% # calculating average and upper and lower bounds (+- SD) for NEP and SRP uptake rates for each temp treatment and each month
  ggplot(aes(x = avgSRP, y = avgNEP, color = month)) +
  geom_point() +
  geom_errorbarh(aes(xmin = lowSRP, xmax = upperSRP)) +
  geom_errorbar(aes(ymin = lowNEP, ymax = upperNEP)) +
  labs(x = expression(PO[4]~Uptake~(ugP~hr^{-1}~gAFDW^{-1})), y = expression(NEP~(mgO[2]~day^{-1}~gAFDW^{-1})))

ggsave("plots/main_analysis/pdef_plot.png", plot = Pdef_plot, height = 5, width = 7, dpi = 300)


# calculating p deficiency index (optimum photosynthetic rate:maximum phosphate uptake velocity) -- units should cancel
# using biomass specific rates
dat_pnld_pdef <- main_dat %>%
  filter(!is.na(SRP_rate)) %>% # filtering out rows with no SRP uptake rate
  filter(exp_ID != 1 & exp_ID != 2) %>% # filtering out march 1st and 15th data since they were performed at a different site
  filter(exp_ID == 2 & temp_C == 7 | exp_ID == 3 & temp_C == 7 | exp_ID == 4 & temp_C == 7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only data from in-situ temperatures in the field
  mutate(P_def_AFDW = NEP_mgO2_d_gAFDW_2/24 * 1000/SRP_rate) %>% # converting NEP to ug O2 per hour per gram AFDW of periphyton and then dividing by the phosphate uptake rate
  group_by(date) %>% # grouping the data by date
  summarise(avg_P_def = mean(P_def_AFDW), lower = mean(P_def_AFDW) - 2 * sd(P_def_AFDW), upper = mean(P_def_AFDW) + 2 * sd(P_def_AFDW))
#View(dat_pnld_pdef)


#################### plotting results
dat_pnld_pdef %>%
  filter(!is.na(lower)) %>% # filtering out rows from 3/15 and 4/21 experiments
  ggplot(aes( x = date, y = avg_P_def)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  labs(x = "Date", y = expression(NEP:P~uptake~(mgO[2]/ugP)))


############## plotting p deficiency index vs. temperature
  main_dat %>%
  filter(!is.na(SRP_rate)) %>% # filtering out rows with no SRP uptake rate
  mutate(P_def_AFDW = NEP_mgO2_d_gAFDW_2/24 * 1000/SRP_rate) %>%
    ggplot(aes(x = temp_C, y = P_def_AFDW)) +
    geom_point() +
    labs(x = "Temp (C)", y = expression(NEP:P~uptake~(mgO[2]/ugP)))


############## now doing same with areal rates
dat_pnld_pdef <- main_dat %>%
  filter(!is.na(SRP_rate)) %>% # filtering out rows with no SRP uptake rate
  filter(exp_ID != 1 & exp_ID != 2) %>% # filtering out march 1st and 15th data since they were performed at a different site
  filter(exp_ID == 2 & temp_C == 7 | exp_ID == 3 & temp_C == 7 | exp_ID == 4 & temp_C == 7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only data from in-situ temperatures in the field
  mutate(P_def_m2 = NEP_mgO2_d_m2/24/SRP_rate) %>% # converting NEP to mg O2 per hour per gram AFDW of periphyton and then dividing by the phosphate uptake rate
  group_by(date) %>% # grouping the data by date
  summarise(avg_P_def = mean(P_def_m2), lower = mean(P_def_m2) - 2 * sd(P_def_m2), upper = mean(P_def_m2) + 2 * sd(P_def_m2))
#View(dat_pnld_pdef)

################ plotting results
dat_pnld_pdef %>%
  filter(!is.na(lower)) %>% # filtering out rows from 3/15 and 4/21 experiments
  ggplot(aes( x = date, y = avg_P_def)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower, ymax = upper))
# units of SRP uptake are per g AFDW of periphyton, not per m2. Thus the P deficiency indices based off areal rates are not comparable to those based off of biomass specific rates and are much higher.

```

All p deficiency index values fall within the extreme P deficiency group as put forth by Pick & Lean 1981.


## Control rates across seasons
### AFDW-normalized
```{r}
# daily rates
metab_trend_AFDW <- dat_long %>%
  filter(exp_ID == 3 & temp_C ==7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only ambient and in-situ temperature data from all experiments except march one due to different biomass methods giving consistently higher data values
  ggplot(aes(x = date, y = rates_AFDW, group = date)) +
  facet_grid(rows = vars(rate_name_AFDW)) +
  geom_boxplot() +
  labs(x = "", y = expression(Metabolic~Rate~(mgO[2]~day^{-1}~gAFDW^{-1}))) +
  theme_bw() +
  theme(legend.position = "None", strip.text.y = element_text(angle = 0)) # removing legend and making facet labels horizontal


metab_trend_AFDW

# hourly rates
metab_trend_AFDW_hr <- dat_long %>%
  filter(exp_ID == 3 & temp_C ==7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only ambient and in-situ temperature data from all experiments except march one due to different biomass methods giving consistently higher data values
  ggplot(aes(x = date, y = rates_AFDW/24, group = date)) +
  facet_grid(rows = vars(rate_name_AFDW)) +
  geom_boxplot() +
  labs(x = "", y = expression(Metabolic~Rate~(mgO[2]~hr^{-1}~gAFDW^{-1}))) +
  theme_bw() +
  theme(legend.position = "None", strip.text.y = element_text(angle = 0)) # removing legend and making facet labels horizontal


metab_trend_AFDW_hr



# simplified plot for plotly
metab_trend_AFDW_plotly <- dat_long %>%
  filter(exp_ID == 3 & temp_C ==7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only ambient and in-situ temperature data from all experiments except march one due to different biomass methods giving consistently higher data values
  ggplot(aes(x = date, y = rates_AFDW/24, group = date, fill = rate_name_AFDW)) +
  facet_grid(rows = vars(rate_name_AFDW)) +
  geom_boxplot() +
  scale_fill_viridis_d(name = "Hourly Rates") +
  theme_bw() +
  theme(legend.position = "None")


p <- plotly::ggplotly(metab_trend_AFDW_plotly)
p


```

### SA normalized
```{r}
# daily rates
metab_trend_sa <- dat_long_m2 %>%
  filter(exp_ID == 3 & temp_C ==7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only ambient and in-situ temperature data from all experiments except march one due to different biomass methods giving consistenly higher data values
  ggplot(aes(x = date, y = rates_m2, group = date)) +
  facet_grid(rows = vars(rate_name_m2)) +
  geom_boxplot() +
  labs(x = "", y = expression(Metabolic~Rate~(mgO[2]~day^{-1}~m^{-2})), fill = "Metabolic Rate") +
  theme_bw() +
  theme(strip.text.y = element_text(angle = 0))


metab_trend_sa


# hourly rates
metab_trend_sa_hr <- dat_long_m2 %>%
  filter(exp_ID == 3 & temp_C ==7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only ambient and in-situ temperature data from all experiments except march one due to different biomass methods giving consistenly higher data values
  ggplot(aes(x = date, y = rates_m2/24, group = date)) +
  facet_grid(rows = vars(rate_name_m2)) +
  geom_boxplot() +
  labs(x = "", y = expression(Metabolic~Rate~(mgO[2]~hr^{-1}~m^{-2})), fill = "Metabolic Rate") +
  theme_bw() +
  theme(strip.text.y = element_text(angle = 0))


metab_trend_sa_hr


# simplified plot for plotly

metab_trend_sa_plotly <- dat_long_m2 %>%
  filter(exp_ID == 3 & temp_C ==7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only ambient and in-situ temperature data from all experiments except march one due to different biomass methods giving consistenly higher data values
  ggplot(aes(x = date, y = rates_m2/24, group = date, fill = rate_name_m2)) +
  facet_grid(rows = vars(rate_name_m2)) +
  geom_boxplot() +
  scale_fill_viridis_d(name = "hourly rate") +
  theme_bw()
 
p <- plotly::ggplotly(metab_trend_sa_plotly)
p

# calculating average areal metabolic rates

control_rates <- dat_long_m2 %>%
  filter(exp_ID == 3 & temp_C ==7 | exp_ID == 5 & temp_C == 14.5 | exp_ID == 6 & temp_C == 19.5 | exp_ID == 7 & temp_C == 15 | exp_ID == 8 & temp_C == 9.5 | exp_ID == 9 & temp_C == 6) %>% # filtering only ambient and in-situ temperature data from all experiments except march one due to different biomass methods giving consistently higher data values
  group_by(rate_name_m2) %>%
  summarise(avg__hourly_rate_m2 = mean(rates_m2)/24)
#View(control_rates)
```

### combining AFDW and SA metabolic trend plots
#### Figure 3
```{r}
metab_trend <- (metab_trend_AFDW + metab_trend_sa)

ggsave("plots/main_analysis/metab_trends.jpg", plot = metab_trend, height = 6, width = 12, dpi = 300)

metab_trend_hr <- (metab_trend_AFDW_hr + metab_trend_sa_hr) +
  plot_annotation(tag_levels = "a")

ggsave("plots/final_paper/Figure3.jpg", plot = metab_trend_hr, height = 6, width = 12, dpi = 300)

```



